### 2019-02-25 docs/manuscript/main.tex
[genesis]

%*******************************************************************************
\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\graphicspath{{../}} %%%REPOSITORY PATH

\title{Strengths and Weaknesses of Recurrent Quantification Analysis 
in the context of Human-Humanoid Interaction}
\author[1,*]{Miguel Xochicale}
\author[1]{Chirs Baber}
\affil[1]{University of Birmingham, 
	School of Engineering, 
	Birmingham, 
	B15 2TT, 
	UK}
\affil[*]{map479@bham.ac.uk}

%\keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
Movement variability is defined as the variations that occur in motor
performance across multiple repetitions of a task and such behaviour is an
inherent feature within and between each persons' movement. Quantifying
movement variability is still an open problem, particularly when traditional
methods in time domain and frequency domain fail to detect tiny modulations in
frequency or phase for time series. We therefore consider methodologies from
nonlinear dynamics such as reconstructed state space (RSS), uniform time-delay
embedding (UTDE), recurrence plots (RPs) and recurrence quantification analysis
(RQA) metrics. These algorithms require time series measured with sensors that
provide well sampled data with little noise. However, for this study, we are
interested in the analysis of data collected through wearable inertial sensors
(IMUs) and its post-processing effects in each of the nonlinear dynamics
methodologies. Twenty right-handed healthy participants (mean and standard
deviation (SD) age of mean=19.8 (SD=1.39)) performed an experiment in
human-humanoid imitation activities (H-H-I). Then IMUs were attached to the
participants and the humanoid robot performed simple vertical and horizontal
arm movements in normal and faster speed. With four window lengths and three
levels of smoothed time series, we found visual differences in the patterns of
RSSs and RPs. Then using metrics of RQA, we find out that the type of movements
and the level of smoothness affects those metrics. In particular, we found that
entropy values from RQA were well distributed and presenting variation in all
the conditions for time series. Hence, we demonstrated the potential of
nonlinear techniques to quantify human movement variability in the context of
H-H-I can enhance the development of better diagnostic tools for various
applications rehabilitation, sport science or for new forms of human-robot
interaction.
%280
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}
Human movement variability is a complex system where not only multiple
joints and limbs are involved for a specific task in a determined environment
but also how we process the external information with all of our available senses 
and use our experiences \cite{davids2003}. 
Recent studies in human motion recognition have revealed the possibility to 
estimate features from lower dimensions signals to distinguish differences between 
styles of pedalling motion \cite{Quintana-Duque2012, Quintana-Duque2016}, 
gait identification \cite{sama2013, frank2010} 
or pattern recognition from biological signals \cite{gomezgarcia2014}.
The lower dimension signals from biological signals are time series 
of $1-$dimension in $\mathbb{R}$ which commonly are 
noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
Hence, we are hypothesising that nonlinear dynamics can be used to 
objectively quantify such human movement variability of lower
dimension signals \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, 
frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
For instance, Bradley et al. 2015 \cite{bradley2015} reviewed methods for
nonlinear time series analysis based on the appropriate estimation of the 
embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to 
reconstruct the state space,
where an $n$-dimensional reconstructed state space using $1-$dimensional 
time series,
can preserve the topological properties of an unknown $M$-dimensional 
state space \cite{takens1981}.
Also, Bradley et al. 2015 \cite{bradley2015} reviewed the use of 
Recurrence Plots (RPs), a graphical representation of a two-dimensional map 
which show black and white dots as recurrences in a given $n$-dimensional system, 
and Recurrence Quantification Analysis (RQA) metrics can pick out important 
directions and statistics in RPs.
In general, RPs and RQA provide a more intuitive meaning of the time series,
for instance, RQA is quantitatively and qualitatively independent of 
embedding dimension (also verified experimentally \cite{iwanski1998}).
However, the estimation of embedding parameters and the selection of the 
right parameters to perform RQA is still an open problem.
Bradley et al. 2015 \cite{bradley2015} pointed out that
there is no general technique that can be used to compute the embedding
parameters since the time series are system-dependent which means that
once computing the values for embedding parameters, these may only work for
one purpose (e.g., prediction) and may not work well for another purpose
(e.g., computing dynamical invariants).
Additionally, the methodologies of nonlinear dynamics for 
computing the embedding parameters e.g., autocorrelation, mutual information, 
and nearest neighbour require data which is well 
sampled and with little noise \cite{garland2016}
or require purely deterministic signals \cite{kantz2003}.
Similarly, these nonlinear analysis methods for computing the embedding 
parameters can break down
with real-world datasets which have generally different length, 
different values of accuracy and precision (rounding errors due to finite 
precision of the measurement apparatus which include frequency 
acquisition \cite{frank2010}),
and data may be contaminated with different sources of noise
\cite{garland2016}.
It is surprising that even with the previous constraints with regard to 
the quality of data, and the problem with the estimation of embedding parameters,
the use of nonlinear dynamics have proven to be helpful to understand and 
characterise dynamics of time series 
\cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010,
gomezgarcia2014, gomezgarcia2014, marwan2011, stergiou2011, bradley2015}.

%Another point to consider with time series analysis using nonlinear dynamics
%is the appropriate use of different post-processing techniques such as 
%interpolation, filtering or normalisation.
%With that in mind, it can be said that there is little research and 
%understanding on the effects for post-processing techniques
%in the interpretation of reconstructed state spaces, RPs and RQA.

For this work, we are interested in exploring and investigating  
the effects of different features of time series 
(e.g. levels of smoothness, window data length,
structures of time series based on types of movements, 
types of sensors, participants and velocities) to estimate 
embedded parameters for uniform time-delay embedding, 
recurrence plots and recurrence quantification analysis.
Hence, we conducted an experiment with twenty right-handed 
healthy participants in the context of human-humanoid imitation 
activities where participants were asked to 
imitate simple arm movements performed by a humanoid robot.
The primary aim of this work is to explore the following questions:
\begin{itemize}
\item What are the effects on three nonlinear methods 
	(reconstructed state space with uniform time-delay embedding, 
	Recurrence Plots, and Recurrence Quantification Analysis), 
	for different embedding parameters, different recurrence thresholds 
	and different characteristics of time series 
	(structure of the signal, window length and smoothness of the signal)?, and 
\item What are the strengths and weaknesses of Recurrence
	Quantification Analysis for the previous conditions
	of nonlinear analysis methods?
\end{itemize}

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Results}
To give insight to the research questions, 
we considered various conditions of time series collected for this work 
(see Experiment section for more details) which are described as follows
\begin{itemize}
\item Three levels of smoothness for the normalised data 
	(sg0zmuv, sg1zmuv and sg2zmuv), computed from two different filter 
	lengths (29 and 159) with the same polynomial degree 
	of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal},
\item four velocities arm movement activities: 
	horizontal normal (HN), horizontal faster (HF), 
	vertical normal (VN), and vertical faster (VF), and
\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
	10-sec (500 samples) and 15-sec (750 samples).
\end{itemize}

%\subsection*{Time series}
After the data collection, raw time series were windowed, normalised and 
smoothed. Then, due to space limitations and to have simple visualisation, 
we only present 10-sec (500 samples) window length time series for 
three participants (p01, p01 and p03) performing horizontal 
arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}). 
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
    	\caption{ 
	{\bf Time series for horizontal and vertical arm movements.}
		(A/D) raw-normalised (sg0zmuv), 
		(B/E) normalised-smoothed 1 (sg1zmuv) and
		(C/F) normalised-smoothed 2 (sg2zmuv).
		Time series are only for three participants 
		($p01$, $p02$, and $p03$) 
		for horizontal and vertical arm movements in normal 
		and faster velocity (HN, HF, VN, VF) 
		with the normalised GyroZ or GyroY axis 
		and with one sensor attached to the participant (HS01) 
		and other sensor attached to the robot (RS01).	
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:ts}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

%\subsection*{Reconstructed State Spaces using Uniform Time-Delay Embedding}
\subsection*{Reconstructed State Spaces}
One of the main challenges of the implementation of 
Uniform Time-Delay Embedding to Reconstructed State Spaces is the 
selection of embedding parameters because each time series is unique in 
terms of its structure (modulation of amplitude, frequency and phase) 
\cite{ frank2010, sama2013, bradley2015}.
With that in mind, the problem for this work is not to compute individual 
embedding parameters for each of the time series but to deal with a 
selecting of two parameters that can represent all the time series. 
Our solution for that problem was to compute a sample mean over all 
values in each of the conditions of the time series 
%of Figs~\ref{fig:ts}
for minimum dimension values and 
%Figs~\ref{fig:caoami} 
for minimum delay values as follows.
%, resulting in an average minimum embedding parameters 
%of ($\overline{m_0}=6$, $\overline{\tau_0}=8$).

\subsubsection*{Minimum embedding parameters}
Minimum embedding parameters were firstly computed 
with False Nearest Neighbour and Average Mutual Information.
%for Uniform Time-Delay Embedding (UTDE),
%Reconstructed State Space (RSS), 
%Recurrence Plots (RPs) and 
%Recurrence Quantification Analysis (RQAs).
Figures \ref{fig:caoami}(A) show that minimum embedding values 
for sensor HS01 appear to show more variations as their 
interquartile range of the box plots are near to 1 
with three exceptions (HF sg0, VN sg0, and VF sg1).
Minimum embedding values appear to be constant for sensor RS01 
as their interquartile range 
in the box plots are near to 0.1 with the exception of 
two axis (HF sg0 and VF sg1).
Additionally, it can be seen in Figs \ref{fig:caoami}(A) a 
decrease of mean values (rhombus) in the box plots
as smoothness of time series increase.
Similarly, the first minimum values of AMI values are 
shown in the box plots of Figs \ref{fig:caoami}(B).
It can be seen that values for HS01 tend to be more spread as the smoothness 
of the time series is increasing 
(see the increase of both mean (rhombus) and interquartile range).
However, AMI values for RS01 do not show such increase in relation with
the increase of smoothness excepting for HF and VF
(Figs \ref{fig:caoami}(B)).

We also computed an overall minimum embedding parameters that represent 
participants, activities, sensors and levels of smoothness, using a sample 
mean of all values in Figs~\ref{fig:caoami}.
The sample mean for the minimum values of $E_{1}(m)$ from Figs \ref{fig:caoami}(A) 
is $\overline{m}_0=6$ 
and the sample mean for minimum values of AMIs from Figs~\ref{fig:caoami}(B)
is $\overline{\tau}_0=8$.
%Hence, we use the average minimum embedding parameters 
%($\overline{m_0}=6$, $\overline{\tau_0}=8$) to compute 
%Reconstructed State Spaces (RSSs), Recurrence Plots (RPs) and
%Recurrence Quantification Analysis (RQA) metrics for human-humanoid activities.
%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/caoami/pdf/fig2}
	\caption{
	{\bf Box plots of minimum embedding parameters for horizontal and vertical arm movements.} 
	Box plots of (A) minimum embedding dimensions 
	and (B) first minimum AMI values for 
	Horizontal Normal (HN), Horizontal Faster (HF),
	Vertical Normal (VN) and Vertical Faster (VF)
	with sensors attached to participants (HS01) and
	sensor attached to robot (RS01).
	Minimum embedding dimensions are for twenty participants 
	($p01$ to $p20$) with three smoothed signals 
	(sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ)
	and window length of 10-sec (500 samples).
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:caoami}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsubsection*{Uniform Time-Delay Embedding}
Uniform time-delay embedding were computed with
embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$) and 
the first three axis of the rotated data of the PCA are shown 
for the reconstructed state spaces in 
Figs~\ref{fig:rss_aHw10} for horizontal arm movements and 
Figs~\ref{fig:rss_aVw10} for vertical arm movements.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig3}
\caption{
	{\bf RSSs for horizontal arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters $m=6$, $\tau=8$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:rss_aHw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig4}
    \caption{
	{\bf RSSs for vertical arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters $m=6$, $\tau=8$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:rss_aVw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

Evidently, it is easy to observe by eye the differences in each of the
trajectories in the reconstructed state spaces 
(Figs~\ref{fig:rss_aHw10}, \ref{fig:rss_aVw10}), 
however one might be not objective when quantifying those differences 
since those observation might vary from person to person.
With that in mind, we tried to objectively quantify those differences 
using euclidean distances between the origin to each of the points in the 
trajectories, however these results created a suspicious metric, specially 
for trajectories which looked very messy.
With that in mind, we computed Recurrence Plots and 
Recurrence Quantification Analysis to objectively quantify 
the differences in each of the cases of the time series.

\subsection*{Recurrences Plots}
Considering the time series of Fig~\ref{fig:ts}, 
Recurrence Plots were computed for horizontal arm movements (Fig~\ref{fig:rp_aH}) and
for vertical arm movements (Fig~\ref{fig:rp_aV}) 
using the average embedding parameters ($m=6$, $\tau=8$) 
and an recurrence threshold of $\epsilon=1$.
With regard to the selection of recurrence threshold,
Marwan et al. \cite{marwan2011} pointed out that choosing an appropriate 
recurrence threshold is crucial to get meaningful representations in RPs, 
however, for this work where quantifying movement variability is our aim,
we do not compute recurrence thresholds as as long as it is able to 
represent the dynamical transitions in each of the time series.

As similar as with the Reconstructed State Spaces, 
the differences in the RPs can be easily noticed by eye 
for different time series (Figs~\ref{fig:rp_aV} and ~\ref{fig:rp_aH}), 
which lead us to apply Recurrence Quantification Analysis 
to have an objective quantification of each of the time series.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aH}
\caption{
	{\bf RPs for horizontal arm movements.}	
	Recurrence plots were computed with 
	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:rp_aH}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aV}
\caption{
	{\bf RPs for vertical arm movements.}	
	Recurrence plots were computed with 
	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:rp_aV}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsection*{Recurrence Quantification Analysis} \label{ch6:rqas}
Considering the RPs for 20 participants performing four activities 
(HN, HF, VN and VF) with sensors attached to the human (HS01) and to the 
humanoid robot (RS01) and with the increase of smoothness of time series 
(sg0zmuvGyroZ, sg1zmuvGyroZ and sg2zmuvGyroZ), 
we hence compute four metrics of RQA metrics (REC, DET, RATIO and ENTR) with 
embedding parameters $m=6$, $\tau=8$ and recurrence threshold $\epsilon=1$.

%\newpage
\subsubsection*{REC values}
It can be seen in the box plots of Figs~\ref{fig:RQABP}(A) that REC values, 
representing the \% of black dots in the RPs, 
are more spread for HN and VN movements (higher interquartile range) 
than HF and VF movements (lower interquartile range) for HS01 sensor. 
In contrast, REC values for RS01 sensor present little variation 
(interquartile range of 0.01).
With regard to the increase of smoothness of time series 
(sg0, sg1 and sg2), REC values present little 
variation as the smoothness is increasing for time series from HS01 
(changes of mean values (rhombus)) while REC values are more affected with 
the smoothness for data from RS01 
(see the incremental changes of mean values (rhombus)).
%See Figs~\ref{fig:rec_aH} and \ref{fig:rec_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual REC values for each participant.

\subsubsection*{DET values}
Figs \ref{fig:RQABP}(B) illustrate DET values, 
representing predictability and organisation of the RPs, 
which change very little (interquartile range is around 0.1) 
for type of movement, level of smoothness or type of sensor.
%See Figs~\ref{fig:det_aH} and \ref{fig:det_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual DET values for each participant.

\subsubsection*{RATIO values}
Figs \ref{fig:RQABP}(C) present RATIO values, representing dynamic transitions, 
for horizontal and vertical movements.
It can be seen that RATIO values for HS01 sensor vary less 
for HN movements (interquartile range around 2)
than HF movements (interquartile range around 5).
%which is a similar behaviour of RATIO values for oRS01 sensor.
It can also be noticed a decrease of variation in RATIO values as the 
smoothness of the time series is increasing (grey rhombus).
%See Figs~\ref{fig:ratio_aH} and \ref{fig:ratio_aV} in appendix 
%\ref{appendix:e:ep} 
%for more details about individual RATIO values for each participant.

\subsubsection*{ENTR values}
Figs \ref{fig:RQABP}(D) show ENTR values, representing the complexity of 
the structure the time series, for both horizontal and vertical movements.
ENTR values for HS01 sensor show more variation 
(interquartile range around 0.5)
than ENTR values for RS01 sensor which appear 
to be more constant (interquartile range 0.1).
It can also be said that the smoothness of time series affects
each of the axis by an increase of mean values (see gray rhombos).
%See Figs~\ref{fig:entr_aH} and \ref{fig:entr_aV} in appendix
%\ref{appendix:e:ep}
%for more details about individual ENTR values for each participant.

%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/pdf/rqa-bp}
    \caption
	[Box plots for RQA values]{
	{\bf Box plots for RQA values.}
	Box plots of (A) REC, (B) DET, (C) RATIO, and (D) ENTR values 
	for 20 participants performing HN, HF, VN and VF movements
	with sensors HS01, RS01 and three smoothed-normalised  
	time series (sg0, sg1 and sg2).
	RQA values were computed with 
	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:RQABP}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

%
%
%\subsection*{RQA metrics with different embedding parameters, 
%	recurrence thresholds, window lengths, levels of smoothness, 
%	and time series structures.}
%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
%determining embedding parameters, their method consisted on creating 3D 
%surfaces with RQA metrics with an increase of embedding parameters 
%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
%fluctuations and gradual changes in the 3D surfaces that provide information 
%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
%created 3D surfaces for visual selection of not only embedding parameters 
%but also recurrence thresholds. Following same methodologies, we explored 
%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
%using 3D surfaces by an unitary increase of the pair embedding 
%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
%(Fig.~\ref{fig:topo_rqas}). We also computed 3D surfaces of RQA metrics 
%for different sensors and different activities 
%(Fig.~\ref{fig:topo_sensoractivities}). RQA metrics are also affected by 
%the window length where for example four window lengths of 100, 250, 500 
%and 750 samples (Fig.~\ref{fig:topo_sensoractivities}). Three level of 
%smoothness were computed for RQA metrics showing smoothed 3D surfaces and 
%the level of smoothness increase (Fig.~\ref{fig:topo_smoothness}).
%Similarly, 3D surfaces of RQA metrics were also computed for three 
%participants (Fig.~\ref{fig:topo_participants}).
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/rqas}
%%\includegraphics[width=1.0\textwidth]{rqas}
%    \caption{
%	{\bf 3D surfaces for RQA metrics.}
%	3D surfaces for REC, DET, RATIO and ENTR values with increasing 
%	pair embedding parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) 
%	and recurrence thresholds (  $ 0.2 \ge \epsilon \le 3 $).
%	RQA metrics values for time series of participant p01 using HS01 sensor, 
%	HN activity and sg0zmuvGyroZ axis and 500 samples window length.
%        R code to reproduce the figure is available from \cite{hwum2018}.
%	}
%\label{fig:topo_rqas}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/sa}
%%\includegraphics[width=1.0\textwidth]{sa}
%    \caption{
%	{\bf 3D surfaces of RQA metrics for sensors and activities.}
%	3D surfaces with increasing embedding parameters and recurrence 
%	thresholds are for HS01 and RS01 sensors of HN, HF, VN and VF 
%	activities. RQA metrics values are for time series of 
%	participant p01 for sensors (HS01 and RS01), 
%	activities (HN, HF, VN and VF) and for 	sg0zmuvGyroZ axis with 
%	500 samples window length. 
%	R code to reproduce the figure is available from \cite{hwum2018}.
%       }
%\label{fig:topo_sensoractivities}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/w}
%%\includegraphics[width=1.0\textwidth]{w}
%    \caption{
%	{\bf 3D surfaces for RQAs metrics with four window lengths.}
%	3D surfaces of RQAs metric values with increasing embedding 
%	parameters and recurrence thresholds are for four window lengths 
%	(w100, w250, w500 and  w750).
%	RQA metrics values are for time series of participant p01 
%	using HS01 sensor, HN activity and sg0zmuvGyroZ axis.
%	R code to reproduce the figure is available from \cite{hwum2018}.
%        }
%\label{fig:topo_sensoractivities}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/s}
%%\includegraphics[width=1.0\textwidth]{s}
%    \caption{
%	{\bf 3D surfaces for RQA metrics with three levels of smoothness.}
%       	3D surfaces of RQAs metric values with increasing embedding 
%	parameters and recurrence thresholds are for three levels of 
%	smoothness (sg0zmuvGyroZ, sg1zmuvGyroZ and sg1zmuvGyroZ).
%	RQA metrics values are for time series of participant p01 using 
%	HS01 sensor, HN activity and 500 samples window length.
%	R code to reproduce the figure is available from \cite{hwum2018}.
% }
%\label{fig:topo_smoothness}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%
%
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/p}
%%\includegraphics[width=1.0\textwidth]{p}
%    \caption{
%	{\bf 3D surfaces for RQA metrics with three participants.}
%       	3D surfaces of RQAs metric values for participants p01, p02 and p03 
%	with increasing embedding parameters and recurrence thresholds.
%	RQA metrics values are for time series of HS01 sensor, 
%	HN activity and 500 samples window length.
%	R code to reproduce the figure is available from \cite{hwum2018}.
% }
%\label{fig:topo_participants}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%
%
%
%
%
%
%
%
%
%
%
%%*******************************************************************************
%%*******************************************************************************
%%*******************************************************************************
%\section*{Discussion}
%%The Discussion should be succinct and must not contain subheadings.
%
%It is evidently that time series from different sources (participants, 
%movements, axis type, window length or levels of smoothness) 
%presents visual differences for embedding parameters and therefore for RRSs. 
%For which, the selection of embedding parameters was our first challenge 
%where we computed embedding parameters for each time series and then computed 
%a sample mean over all time series in order to get two embedding parameters 
%to compute all RSSs with its corresponded type of movement. Then we found 
%that the quantification of variability with regard to the shape of the 
%trajectories in RSSs requires more investigation since our original 
%proposed method base on euclidean metric failed to quantify those 
%trajectories. Specially, for trajectories which were not well unfolded. 
%With that in mind, we proceed to take advantage of four RQA metrics 
%(REC, DET, RATIO and ENTR) in order to avoid any subjective interpretations 
%or personal bias with regard to the evolution of the trajectories in RSSs.
%
%
%\subsection*{RQA metrics with fixed parameters}
%Considering that RQA metrics were computed with fixed embedding parameters 
%($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found 
%the following. REC values, which represents the \% of black points in the RPs, 
%were more affected with and increase in normal speed movements (HN and VN) 
%than faster movements (HF and VF) for the sensor attached to the participants 
%(HS01). Such decrease of REC values from normal speed to faster speed 
%movements is also presented in data from sensor attached to the robot (RS01), 
%and little can be said with regard to the dynamics of the time series coming 
%from RS01. Similarly, DET values, representing predictability and 
%organisation in the RPs, present little variation in the any of the time 
%series where little can be said. In contrast, RATIO values, which represent 
%dynamic transitions, were more variable for faster movements (HF and VF) 
%than normal speed movements (HN and VN) with sensors attached to the 
%participants (HS01). For data coming from sensors attached to the robot 
%(RS01), RATIO values from horizontal movements (HN, HF) appear to vary 
%more than values coming from vertical movmentes (VN, VF).
%With that, it can be said that RATIO values can represent a bit better
%than REC or DET metrics for the variability of imitation activities in 
%each of the conditions for time series.
%In the same way, ENTR values for HN were higher than values for HF
%and ENTR values varied more for sensor attached to participants 
%than ENTR values for sensors of the robot. It is evidently that 
%the higher the entropy the more complex the dynamics are, 
%however, ENTR values for HN appear a bit higher than HF values, 
%for which we believe this happens because of the structure the time series
%which appear more complex for HN than  HF movements which presented a 
%more consistence repetition.
%
%We observed that some RQA metrics are affected by the smoothness of data.
%For which, we also explored the effect of smoothness of raw-normalised data 
%where, for example, REC and DET values were not affected by the smoothness 
%of data since these seemed to be constants. However, for RATIO values, 
%the effect of smoothness can be noticed with a slight decrease of amplitude 
%in any of the time series conditions which is also presented with ENTR values.
%
%
%\subsection*{RQA metrics with different parameters}
%Patterns in RPs and metrics for RQA are independent of embedding dimension 
%parameters \cite{iwanski1998}, however, that is not the case when using 
%different recurrence thresholds. Such changes of recurrence threshold values 
%can modify the patterns in RPs and therefore the values of RQA metrics.
%We therefore computed 3D surfaces to explore the sensibility and robustness of 
%embedding parameters and recurrence threshold in RQA  metrics. Following the 
%same methodology of computing 3D surfaces, we also considered variation of 
%window length size to present RQA metrics dependencies with embedding 
%parameters, recurrence thresholds and window length size.
%
%
%
%\section*{Conclusions}
%We conclude that using a different level of smoothness for time series help us 
%to visualise and to quantify the variation of movements between participants 
%using RSSs, RPs and RQA. Also, it is important to mention that some RQA's 
%metrics (e.g. DET and  ENTR) are more robust to the effect of smoothness 
%of time series.
%
%Althought our our experiment is limited to twenty healthy right-handed 
%participants of a range age of mean 19.8 SD=1.39, RQA metrics can quantify 
%human movement variability. With that in mind, we conclude that 
%quantification of human-humanoid imitation activities is possible for 
%participants of different ages, state of health and anthropomorphic features.
%
%
%%
%%Such understanding and measurement of movement variability using
%%cheap wearable inertial sensors lead us to have a more intuitive selection of parameters
%%to reconstruct the state spaces and to create meaningful interpretations
%%of the recurrence plots and the results of the metrics with recurrence quantification 
%%analysis. 
%%
%
%In general, activity type, window length and structure of the time series 
%affects the values of the metrics of RQA for which certain RQA metrics 
%are better to describe determined type of movement.
%Using determined RQA metrics depends on what one want to quantify, for instance, 
%one can find predicability, organisation of the RPs, dynamics transitions, 
%or complexity and determinism.
%
%
%%
%%However, we believe that further investigation is required to find the 
%%right balance between the level of smoothness of the signal
%%and defining what is the aim of RQA where 
%%its representations using RSS, RP and RQA.
%%Specially, where the level of smoothness does not affect 
%%the variation of each of the movements quantification. 
%%%We can also conclude that finding the right balance between 
%%%smoothness and the raw data to capture movement variability is 
%%%a still a problem that has many avenues for exploration.
%%
%
%Similarly, such differences in time series created differences in each of the 
%RQA metrics, for instance, RATIO and ENTR are helpful to distinguish 
%differences in any of the categories of the time series (sensor, activity, 
%level of smoothness and number of participant), however for certain time series 
%(data from the sensor attached to the robot) seemed to have little variations 
%between each of the participants. The latter phenomena was in a way evidently
%as robot degrees of freedom did not allow it to move with a wide range of 
%variability. 
%
%
%
%\section*{Future Work}
%
%\subsection*{Inertial Sensors}
%To have more fundamental understating of nature of signals collected through 
%inertial sensors in the context of human-robot interaction, we are considering 
%to apply derivates to the acceleration data. We can then explore the jerkiness 
%of movements and therefore the nature of arm movements which typically have 
%minimum jerk \cite{flash1985}, its relationship with different body 
%parts \cite{devries1982, mori2012} or the application of higher derivatives 
%of displacement with respect time such as snap, crackle and 
%pop \cite{eager2016}.
%
%\subsection*{RQA}
%Having presented our results with RQA metrics, we believe that further 
%investigation is required to have more robust metrics. For example, 
%Marwan et al. \cite{marwan2007, marwan2015} reviewed different aspects to 
%compute RPs using different criteria for neighbours, different norms 
%($L_{1-norm}$, $L_{2-norm}$, or $L_{\infty-norm}$ ) or different methods 
%to select the recurrence threshold $\epsilon$, such as using certain 
%percentage of the signal \cite{letellier2006}, the amount of noise or 
%using a factor based on the standard deviation of the observational noise 
%among many others \cite{marwan2007}.
%
%
%
%
%
%
%
%%*******************************************************************************
%%*******************************************************************************
%%*******************************************************************************
%\section*{Materials and methods}
%\subsection*{State Space Reconstruction}
%The method of state space reconstruction was originally proposed by 
%\cite{packard1980} and formalised by \cite{takens1981}. Since then, different 
%investigations and disciplines have benefited from the use of the method of 
%state space reconstruction \cite{aguirre2009, stergiou2011, frank2010, 
%sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
%based on uniform time-delay embedding methodology which is a simple 
%matrix implementation that can reconstruct an unknown $d-$dimensional 
%manifold $M$ from a scalar time series (e.g. one-dimensional 
%time series in $\mathbb{R}$). A manifold, in this context, is a 
%multidimensional curved surface within a space (e.g. a saddle) 
%\cite{guastello-gregson2011}.
%
%The use of a scalar time series is the main advantage of the method of state 
%space reconstruction which in essence preserve dynamic invariants such as 
%correlation dimension, fractal dimension, Lyaponov exponents, Kolmogorov-Sinai 
%entropy and detrended fluctuation analysis \cite{bradley2015, 
%Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
%However, selecting appropriate embedding parameters which are sued to apply the 
%state space reconstruction is still an open challenge for which we present 
%introductions for the methodologies to compute such embedding parameters 
%With that in mind, in the following subsections, we describe in more detail 
%the state space reconstruction theorem (RSSs), uniform time-delay embedding 
%theorem (UTDE), false nearest neighbours (FNN), average mutual 
%information (AMI) and other methodologies for state space reconstruction.
%
%\subsection*{State Space Reconstruction Theorem}
%Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
%uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:ssr}
%  s(t)=f^t [s(0)],
%\end{equation}
%%%********************************[EQUATION]************************************
%where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
%$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
%unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
%function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
%iteration of $f$ that corresponds to an initial position 
%$s(0) \in M $ \cite{takens1981}.
%%$f$ is a smooth dynamical system, where smooth means a 
%%continuously differentiable system
%%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
%Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
%can be obtained with
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:measurement}
%  x(t)=h[s(t)],
%\end{equation}
%%%********************************[EQUATION]************************************
%where $h$ is a function, $h: M \rightarrow \mathbb{R}$, defined on the trajectory $s(t)$.
%Reconstructed state space can then be described as an $n-$dimensional state 
%space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
%$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
%is the uniform time-delay embedding with a dimension embedding $m$
%and delay embedding $\tau$ and
%$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
%of dimensionality (e.g. Principal Component Analysis, 
%Singular Value Decomposition, etc) being $n \leq m$.
%With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
%defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
%$\boldsymbol{X}(t) = \Phi(s(t))$,
%where $\Phi$ is a diffeomorphic map \cite{takens1981}
%whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
%dimension of $M$ \cite{garland2016}.
%% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
%% is called diffeomorphic if it is one-to-one correspondence and its inverse
%% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
%Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
%space then a composition of functions represented with $F^t$ is induced on the
%reconstructed state space determined:
%%by true smooth dynamical system $f^t$ and $\Phi$:
%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:st}
%  \boldsymbol{X}(t)=F^t [\boldsymbol{X}(0)] = \Phi \circ f^t \circ \Phi ^{-1}[\boldsymbol{X}(0)].
%\end{equation}
%%%********************************[EQUATION]************************************
%With this in mind, an embedding is defined as "a smooth one-to-one 
%coordinate transformation with a smooth inverse" and the total reconstruction 
%map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
%Fig~\ref{fig:ssr} illustrates the state space reconstruction.
%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/rss/casdagli1991/figure1/rss_v04}
%%\includegraphics[width=1.0\textwidth]{rss_v04}
%    \caption{
%	{\bf State space reconstruction methodology.}
%	State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$
%	where $f^t$ is the true dynamical system, $s(t)$ indicates the state,
%    	$s$, at time, $t$,  and $h[ ]$ the measurement function. 
%    	The time-delay embedding represented as the $\Phi$, maps the original
%    	$d-$dimensional state $s(t)$ into the $m-$dimensional uniform 
%	time-delay embedding matrix $\boldsymbol{X}(t)$.
%    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into 
%	a new state $y(t)$ of dimensions $n < m$.
%    	(A) $M-$dimensional manifold representing the state space 
%	(e.g. Lorenz system);
%    	(B) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
%   	(C) $m-$dimensional reconstructed state space with 
%    	\texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and 
%    	(D) $y(t)$ is the $n-$dimensional reconstructed state space.
%	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $
%	where $\Phi$ is the delay reconstruction map and 
%	$\Psi$ is the coordinate transformation map.
%	This figure is adapted from 
%   	\cite{Quintana-Duque2012, casdagli1991, uzal2011}
%	and R code to reproduce it is available from \cite{hwum2018}.
%    }
%    \label{fig:ssr}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%
%
%
%\subsection*{Uniform Time-Delay Embedding (UTDE)}\label{sec:utimedelayembedding}
%Frank et al. and Sama et al. refer to the state space reconstruction
%as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}. 
%However, we consider the term "uniform time-delay embedding"
%as more descriptive and appropriate terminology for our work.
%
%The uniform time-delay embedding is represented as a matrix of uniform delayed 
%copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is 
%the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the 
%samples of $\{ \boldsymbol{x}_n \}$.
%$\{ \boldsymbol{x}_n \}_{n=1}^N$ has a sample rate of $T$.
%The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$
%and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$
%where $i$ goes from $0,1, \dots, (m-1)$.
%Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains
%information of unobserved state variables and encapsulates the information of
%the delayed copies of the available time series in the uniform time-delay 
%embedding matrix $\boldsymbol{X}^{m}_{\tau}$, 
%$\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:tde}
%\boldsymbol{X}^{m}_{\tau}  =
%\begin{pmatrix}
%\boldsymbol{ \tilde{x} }_n \\
%\boldsymbol{ \tilde{x} }_{n-\tau} \\
%\boldsymbol{ \tilde{x} }_{n-2\tau} \\
%\vdots \\
%\boldsymbol{ \tilde{x} }_{n- (m-1) \tau} \\
%\end{pmatrix}^\intercal, 
%\end{equation}
%%%********************************[EQUATION]************************************
%where $m$ is the embedding dimension, $\tau$ is the embedding delay and
%$ ^\intercal$ denotes the transpose.
%$m$ and $\tau$ are known as embedding parameters.
%The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by
%$N-(m-1)\tau$ rows and $m$ columns and 
%$N-(m-1)\tau$ defines the length of each delayed copy 
%of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
%%For further details and explicit examples of uniform time-delay 
%%embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.
%
%
%
%
%\subsection*{Estimation of Embedding Parameters}
%The estimation of the embedding parameters ($m$ and $\tau$) 
%is a fundamental step for the state space reconstruction with the use
%of uniform time-delay embedding method.
%With this in mind, we review two of the most common algorithms,
%which will be used in our work, to compute the embedding
%parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%*******************************************************************************
%\subsubsection*{False Nearest Neighbours}
%To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992}
%used the method of false neighbours which can be understood as follows:
%on the one hand, when the embedding dimension is too small to unfold the 
%attractor "not all points that lie close each other will be neighbours and 
%some points appear as neighbours because of the attractor has been projected 
%down into an smaller space", on the other hand, when increasing the embedding 
%dimension "points that are near to each other in the sufficient
%embedding dimension should remain close as the dimension increase from $m$ 
%to $m+1$ \cite{krakovska2015}".
%From a mathematical point of view, the state space reconstruction theorem is 
%done when the attractor is unfolded with either the minimum embedding 
%dimension, $m_0$, or any other embedding dimension value where 
%$m \ge m_0$ \cite{kennel1992}. On the contrary, any large value of $m_0$ 
%leads to excessive computations \cite{bradley2015}. With this in mind, 
%Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method 
%where only the time-series and one delay embedding value are necessary 
%to select the minimum embedding dimension. Cao's algorithm is based 
%on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:e}
%  \begin{aligned}
%	E(m) &= \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau} a(i,m) \\
%	 &=
%       \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
%       \frac{ || \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) || }
%            { || \boldsymbol{X}_i(m) - \boldsymbol{X}_{n(i,m)}(m) ||  }
%  \end{aligned}
%\end{equation}
%%%********************************[EQUATION]************************************
%where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay
%embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
%From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ 
%and $\tau$ for which $E_1(m)$ is defined as
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:e1}
%E_1(m) = \frac{ E(m+1) } { E(m)}.
%\end{equation}
%%%********************************[EQUATION]************************************
%$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$
%in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
%As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater
%than some $m_0$, if the time series comes from a multidimensional state space
%then $m_0 + 1$ is the minimum dimension".
%Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from
%stochastic signals. $E_2(m)$ is defined as
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:e2}
%E_2(m) = \frac{ E^* (m+1) } { E^*(m)},
%\end{equation}
%%%********************************[EQUATION]************************************
%where
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:ee}
%E^*(m) = \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
%|  \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) |.
%\end{equation}
%%%********************************[EQUATION]************************************
%For instance, when the signal comes from random noise (values that are 
%independent from each other), all $E_2(m)$ values are approximately equal 
%to 1 (e.g. $E_2(m) \approx 1$). However, for deterministic data $E_2(m)$ is 
%not constant for all $m$ (e.g. $E_2(m) \neq 1$).
%
%As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
%series: the solution for the $x$ variable of the Lorenz system 
%(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
%and a variance of one (Fig~\ref{fig:e1e2}F).
%We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
%The $E_1(m)$ values for the chaotic time series appear to be constant
%after the dimension is equal to six.
%The determination of six is given that any value of $m$ can be used as they 
%are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
%$E_2(m)$ values, for chaotic time series, are different to one 
%(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
%chaotic time series the minimum embedding dimension the time series 
%comes from a deterministic signal. With regard to the noise time series,  
%$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
%which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
%$E_1(m)$ values then indicate the minimum embedding dimension of the 
%noisy time series is thirteen, however all of the $E_2(m)$ values are 
%approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
%concluded that noise time series is a stochastic signal.
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/falseNearestNeighbours/cao97/fig_v04}
%%\includegraphics[width=1.0\textwidth]{fig_v04}
%    \caption{
%	{\bf Minimum dimension embedding values with Cao's method.} 
%	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
%	with variations of $\tau$ values from one to twenty
%	for (E) chaotic and (F) random time series.
%	R code to reproduce the figure is available from \cite{hwum2018}.
%        }
%    \label{fig:e1e2}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%
%It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
%computed but also a variation of $\tau$ from 1 to 20 is presented. 
%The purpose of such variation for $\tau$ is to show its independence with
%regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
%(Fig~\ref{fig:e1e2}A,B,C, and D). 
%However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
%a new threshold where $m$ values appear to be constant in $E_1 (m)$.
%In the case of the given examples and reported results, we defined such 
%threshold as 0.05. Further investigation is required for the selection of the 
%threshold in the $E_1(m)$, as the selection of the threshold in this work is 
%base on no particular method but visual inspection.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%*******************************************************************************
%\subsubsection*{Average Mutual Information}
%When selecting the delay dimension parameter, $\tau$, one can consider the 
%following two cases:
%(i) when $\tau$ is too small, the elements of time-delay embedding will be along
%the bisectrix of the phase space and the reconstruction is generally not 
%satisfactory, 
%(ii) on the contraty, when $\tau$ is too large the elements of the uniform 
%time-delay embedding will become spread and uncorrelated which makes 
%recovering the underlying attractor more difficult if not 
%impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
%With regard to the algorithms to compute $\tau$, 
%Emrani et al. \cite{emrani2014a}, for instance, used the autocorrelation 
%function in which the first zero crossing is considered as the minimum delay 
%embedding parameter. However, the autocorrelation function is a linear 
%statistic over which the Average Mutual Information (AMI) algorithm is 
%preferred because the AMI takes into account the nonlinear dynamical 
%correlations \cite{afraser1986,krakovska2015}. With this in mind, the AMI 
%algorithm is described below to estimate the minimum delay embedding 
%parameter, \texorpdfstring{$\tau_0$}{T}.
%
%To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated
%and then a probability distribution of data is computed \cite{kantz2003}.
%AMI is therefore denoted by $I(\tau)$ which is the average mutual 
%information between the original time series, $x(n)$, and the delayed 
%time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}. 
%AMI is defined by
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:ami}
%I(\tau) = \sum_{i,j}^N p_{ij} log_2 \frac{ p_{ij} }{ p_i p_j }.
%\end{equation}
%%%********************************[EQUATION]************************************
%Probabilities are defined as follows:
%$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of 
%the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside 
%the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability 
%that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$. The AMI is measured 
%in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
%For small $\tau$, AMI will be large and it will then decrease more or 
%less rapidly. As $\tau$ increase and goes to a large limit, 
%$x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is 
%factorised as $p_ip_j$ for which AMI is close to zero. Then, in order 
%to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ 
%where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or,
%where the redundancy is the least" \cite{kantz2003}.
%
%For example, we compute the AMI for two time series:
%A) the $x$ solution of the Lorenz system, and
%B) a noise time series using a normal distribution with mean zero and 
%standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
%concluded that the amount of knowledge for any noise time series is zero 
%for which the first minimum embedding parameter is $\tau_0=1$. 
%On the contrary, the first minimum of the AMI for the chaotic time series 
%is $\tau_0=17$ which is the value that maximize the independence 
%between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
%\cite{bradley2015}.
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/averageMutualInformation/example/ami_plots/figure_v04}
%%\includegraphics[width=1.0\textwidth]{figure_v04}
%    \caption{
%	{\bf Minimum delay embedding values with AMI's method.} 
%    	(A, B) AMI values where its first minimum value in the curve
%	is the minimum time delay embedding ($\tau_0$), 
%	for (C) a chaotic and (D) noise time series.
%	R code to reproduce the figure is available from \cite{hwum2018}.
%        }
%    \label{fig:amis}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%Similarly as Cao's algorithm negatives, AMI's algorithm is not an
%exception for negatives, which are worthwhile to mention for further 
%investigations.
%For instance, (i) is not clear why the choose of the first minimum of the AMI 
%is the minimum delay embedding parameter \cite{kantz2003} and 
%(ii) the probability distribution of the AMI function is computed
%with the use of histograms which depends on a heuristic choice of number of bins
%for which AMI depends on partitioning \cite{garcia2005e71}.
%
%
%
%
%
%
%%*******************************************************************************
%\section*{Recurrence Quantification}\label{sec:recurrence-quantification}
%\subsection*{Recurrence Plots}
%Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences 
%in conservative systems, however such discovery was not put into practice 
%until the development of faster computers \cite{marwan2007},
%for which Eckmann et al. \cite{eckmann1987} in 1987 introduced a method
%where recurrences in the dynamics of a system can be visualised using 
%Recurrence Plots. The intention of Eckmann et al. \cite{eckmann1987} was to 
%propose a tool, called Recurrence Plot (RP), that provides insights into 
%high-dimensional dynamical systems where trajectories are very difficult to 
%visualise. Therefore, "RP helps us to investigate the 
%$m-$dimensional phase space trajectories through a two-dimensional 
%representation of its recurrences" \cite{marwan2015}.
%Similarly, Marwan et al. \cite{marwan2015} pointed out that additionally 
%to the methodologies of the state space reconstruction and other dynamic 
%invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, 
%the recurrences of the trajectories in the phase space can provide 
%important clues to characterise natural process that present, for
%instance, periodicities (as Milankovitch cycles) or irregular cycles 
%(as El Ni\~no Southern Oscillation). 
%Such recurrences can not only be presented visually using Recurrence Plots (RP) 
%but also be quantified with Recurrence Quantification metrics, which leads 
%to applications of these in various areas such as economy, physiology, 
%neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.
%
%For the creation of a recurrence plot based on time series 
%$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
%reconstruction with uniform time-delay embedding 
%$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
%\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
%where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
%and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
%The recurrence plot is therefore a two-dimensional $N \times N$ square 
%matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
%whenever $X(i)$ is sufficiently close to $X(j)$: 
%%%********************************[EQUATION]************************************
%\begin{equation}
%\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
%\end{equation}
%%%********************************[EQUATION]************************************
%where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
%$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
%function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
%(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
%RP is also characterised with a line of identity (LOI) which is a  
%diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/rqa/figure11/v02/figure11-v02}
%%\includegraphics[width=1.0\textwidth]{figure11-v02}
%    \caption{
%	{\bf Recurrence Plots.} 
%	(A) State space of the Lorenz system with controlling parameters 
%	($\rho=28, \sigma=10, \beta=8/3$). 
%	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
%	(black circle) of a given point at $i$ is a recurrent point and is 
%	represented as a black dot in the recurrence plot at location 
%	$(i, j)$ or white otherwise.
%	(B) Recurrence plot using the 
%	three components of the Lorenz system and the RP with no embeddings 
%	and threshold $\epsilon=5$.
%	This figure is adapted from \cite{marwan2015} and R code to reproduce 
%	it is available from \cite{hwum2018}.
%	}
%    \label{fig:mrp}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%
%
%
%
%
%%*******************************************************************************
%%*******************************************************************************
%%*******************************************************************************
%\subsection*{Structures of Recurrence Plots}
%Pattern formations in the RPs can be designated either 
%as topology for large-scale patterns or texture for small-scale patterns.
%In the case of topology, the following pattern formations are commonly presented:
%(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
%uniformly distributed noise (Fig~\ref{fig:rp2}A), 
%(ii) periodic and quasi-periodic systems where diagonal lines and 
%checkerboard structures represent oscillating systems, e.g., sinusoidal 
%signals (Fig~\ref{fig:rp2}B), 
%(iii) drift where paling or darkening recurrence points away from 
%the LOI is caused by drifting systems, 
%e.g., logistic map (Fig~\ref{fig:rp2}C), and
%(iv) disrupted where recurrence points are presented white areas or 
%bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
%(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
%Texture patterns in RPs can be categorised as:
%(i) single or isolated recurrence points that represent rare occurring states, 
%and do not persist for any time or fluctuate heavily,
%(ii) dots forming diagonal lines where the length of the small-scale parallel 
%lines in the diagonal are related to the ratio of determinism or predictability 
%in the dynamics of the system, and
%(iii) dots forming vertical and horizontal lines where the length of the 
%lines represent a time length where a state does not change or change very 
%slowly and these patterns formation represent discontinuities in the signal, 
%and (iv) dots clustering to inscribe rectangular regions which are related 
%to laminar states or singularities \cite{marwan2015}.
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/rqa/figure12/v02/figure12-v02}
%%\includegraphics[width=1.0\textwidth]{figure12-v02}
%    \caption{
%	{\bf Patterns in Recurrence Plots.} 
%	Time-series with its respective recurrence plot for:
%	(A) uniformly distributed noise,
%	(B) super-positionet harmonic oscillation 
%	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
%	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
%	with a linearly increase term ($0.01*i$), and
%	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
%	This figure is adapted from \cite{marwan2015} and R code 
%	to reproduce it is available from \cite{hwum2018}.
%	}
%    \label{fig:rp2}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%
%Although, each of the previous pattern descriptions of the structures in the 
%RP offer an idea of the characteristics of dynamical systems, 
%these might be misinterpreted and conclusions might tend to be subjective 
%as these require the interpretation of a particular researcher(s).
%Because of that, recurrence quantification analyis (RQA) offer objective 
%methodologies to quantify such visual characteristics of previous 
%recurrent pattern structures in the RP \cite{zbilut1992}.
%
%%*******************************************************************************
%\subsection*{Recurrence Quantifications Analysis (RQA)}
%Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
%the density of recurrence points in RPs, then histograms of lengths for 
%diagonal lines in RPs were studied by \cite{trulla1996} which were the 
%introduction to the term recurrence quantification analysis (RQA) 
%\cite{marwan2008}. RQA has been applied in many fields such as life science, 
%engineering, physics, and others \cite{marwan2008}. Particularly in human 
%movement to investigate noise and complexity of postural control 
%\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
%coordination \cite{duran2017}. The success of RQA is not only due to its 
%simple algorithmic implementation but also to its capacity to detect tiny 
%modulations in frequency or phase which are not detectable using standard 
%methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
%RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
%dimension which is verified experimentally by \cite{iwanski1998}.
%RQA metrics comprehend percentage of recurrence, percentage of determinism, 
%ratio, Shannon entropy of the frequency distributions of the line lengths,
%maximal line length and divergence, trend and laminariy 
%\cite{marwan2007, marwan2015}. For this work, we considered only four 
%RQA metrics, due to its consistency with our preliminary experiments, 
%which are described below. Such metrics are computed the nonlinearTseries 
%R package \cite{nonlinearTseries2016}.
%
%\subsubsection*{REC values}
%The percentage of recurrence (REC) is defined as
%%%********************************[EQUATION]************************************
%\begin{equation}
%REC(\epsilon,N) = 
%		\frac{1}{N^2 - N} \sum^{N}_{i \neq j = 1} 
%		\mathbf{R}^{m}_{i,j}(\epsilon),
%\end{equation}
%%%********************************[EQUATION]************************************
%which enumerates the black dots in the RP excluding the line of identity.
%REC is a measure of the relative density of recurrence points in the sparse 
%matrix \cite{marwan2015}.
%%REC is computed as follow with the nonlinearTseries package 
%% \cite{nonlinearTseries} 
%%  hist = getHistograms(neighs, ntakens, lmin, vmin)
%%  # calculate the number of recurrence points from the recurrence rate. The
%%  # recurrence rate counts the number of points at every distance in a concrete
%%  # side of the main diagonal.
%%  # Thus, sum all points for all distances, multiply by 2 (count both sides) and
%%  # add the maindiagonal
%%  numberRecurrencePoints = sum(hist$recurrenceHist) + ntakens
%%  # calculate the recurrence rate dividing the number of recurrent points at a
%%  # given distance by all points that could be at that distance
%%  recurrence_rate_vector = 
%%  hist$recurrenceHist[1:(ntakens - 1)] / ((ntakens - 1):1)
%%  # percentage of recurrent points
%%  REC = (numberRecurrencePoints) / ntakens ^ 2
%
%\subsubsection*{DET values} 
%The percent determinism (DET) is defined as the fraction of recurrence points
%that form diagonal lines and it is determined by
%%%********************************[EQUATION]************************************
%\begin{equation}
%DET=
%	\frac{\sum^{N}_{l=d_{min}} l H_D{l} }
%	     {\sum^{N}_{i,j=1} \mathbf{R}_{i,j}(\epsilon) },
%\end{equation}
%%%********************************[EQUATION]************************************
%where 
%%%********************************[EQUATION]************************************
%\begin{equation}
%H_D(l) = 
%	\sum^{N}_{i,j=1} 
%	(1- \mathbf{R}_{i-1,j-1}(\epsilon) ) 
%	(1- \mathbf{R}_{i+l,j+l}(\epsilon) ) 
%	\prod^{l-1}_{k=0}  \mathbf{R}_{i+k,j+k}(\epsilon)
%\end{equation}
%%%********************************[EQUATION]************************************
%is the histogram of the lengths of the diagonal structures in the RP.
%DET can be interpreted as the predictability of the system for periodic signals 
%which, in essence, have longer diagonal lines than the short diagonals lines
%for chaotic signals or absent diagonal lines for stochastic signals 
%\cite{marwan2007, marwan2015}. Similarly, DET is considered as a measurement for 
%the organisation of points in RPs  \cite{iwanski1998}. 
%%percent determinism (DET) is computed as follow with the nonlinearTseries 
%%package \cite{nonlinearTseries}  
%% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
%%                                       lmin = 2, lDiagonalHistogram,
%%                                       recurrence_rate_vector, maxDistanceMD) {
%%  #begin parameter computations
%%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
%%  DET = num / numberRecurrencePoints
%
%
%\subsubsection*{RATIO values}
%RATIO is defined as the ratio between DET and REC and it is calculated from 
%the frequency distributions of the lengths of the diagonal lines.
%RATIO is useful to discover dynamic transitions \cite{marwan2015}.
%%  diagP = calculateDiagonalParameters(
%%    ntakens, numberRecurrencePoints, lmin, hist$diagonalHist,
%%    recurrence_rate_vector, maxDistanceMD
%%  )
%% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
%%                                       lmin = 2, lDiagonalHistogram,
%%                                       recurrence_rate_vector, maxDistanceMD) {
%%  #begin parameter computations
%%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
%%  DET = num / numberRecurrencePoints
%% 
%%
%%    RATIO = diagP$DET / REC
%%
%
%\subsubsection*{ENT values}
%ENT is the Shannon entropy of the frequency distribution of the diagonal 
%line lengths and it is defined as
%%%********************************[EQUATION]************************************
%\begin{equation}
%ENT= - \sum^{N}_{l=d_{min}} p(l) ln p(l) \quad with \quad p(l)=\frac{ H_D(l) }{ \sum^{N}_{ l=d_{min} } H_D(l) }.
%\end{equation}
%%%********************************[EQUATION]************************************
%ENT reflects the complexity of the deterministic structure in the system.
%For instance, for uncorrelated noise or oscillations, 
%the value of ENT is rather small and indicates low complexity of the system,
%therefore "the higher the ENT is the more complex the dynamics are" 
%\cite{marwan2007, marwan2015}.
%%#'  \item \emph{ENTR}: Shannon entropy of the diagonal line lengths distribution
%%
%%calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
%%                                       lmin = 2, lDiagonalHistogram,
%%                                       recurrence_rate_vector, maxDistanceMD) {
%%
%%  pl = lDiagonalHistogram / sum(lDiagonalHistogram)
%%  diff_0 = which(pl > 0)
%%  ENTR = -sum(pl[diff_0] * log(pl[diff_0]))
%
% 
%\subsection*{Sensitivity and robustness of RPs and RQA.}
%RP and RQA are a very young field in nonlinear dynamics and many questions 
%are still open, for instance, different parameters for window length size 
%of the time series, embedding parameters or recurrence threshold can 
%generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.
%
%The selection of recurrence threshold, $\epsilon$, can depend on the system 
%that is analysed. For instance, when studying dynamical invariants $\epsilon$ 
%require to be very small, for trajectory reconstruction $\epsilon$ requires 
%to have a large thresholds or when studying dynamical transition 
%there is little importance about the selection of the threshold 
%\cite{marwan2011}. Other criteria for the selection of $\epsilon$ is that 
%the recurrence threshold  should be five times larger 
%than the standard deviation of the observational noise
%or the use of diagonal structures within the RP is suggested in order
%to find the optimal recurrence threshold for (quasi-)periodic process 
%\cite{marwan2011}.
%
%Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
%of choosing the right embedding parameters to perform RQA for which 
%many experiments have to be performed using different parameters in order 
%to have a better intuition of the nature of the time series and how 
%this is represented by using RQA.
%
%With that in mind, this work explores the sensitivity and robustness 
%of the window size of time series, embedding parameters for RSS with UTDE 
%and recurrence threshold for RP and RQA in order to gain a better insight 
%into the underlying time series collected from inertial sensors in the 
%context of human-humanoid imitation activities.
%
%
%
%
%
%
%
%%*******************************************************************************
%%*******************************************************************************
%%*******************************************************************************
%\section*{Experiment} \label{sec:experiment}
%We conducted an experiment in the context of human-humanoid imitation (HHI) 
%activities where participants were asked to imitate simple horizontal and 
%vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
%Such simple movements were repeated ten times for the participant 
%who copied NAO's arm movements in a face-to-face imitation activity.
%Also, wearable inertial measurement unit (IMU) sensors were attached 
%to the right hand of the participant and to the left hand of the robot 
%(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
%IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
%accelerometer, gyroscope and magnetometer sensors and quaternions 
%\cite{Comotti2014}. 
%%A further description of the NeMEMSi IMU sensors 
%%is given in \nameref{S1_AppendixB}.
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%  \centering
%\includegraphics[width=1.0\textwidth]{figures/experiment/hri/hri-v01}
%%\includegraphics[width=1.0\textwidth]{hri-v01}
%    \caption{
%	{\bf Human-humanoid imitation activities.} 
%    		Face-to-face human-humanoid imitation (HHI) activities for 
%		(A) HHI of horizontal arm movement, 
%		(B) Humanoid horizontal arm movement,
%		(C) HHI of vertical arm movement, and 
%		(D) Humanoid vertical arm movement.
%        }
%    \label{fig:hri}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%
%\subsection*{Participants}
%Twenty-three participants,
%from now on defined as $pN$ where $N$ is the number of participant, were 
%invited to do the experiment. However, data for three participants were 
%not used because the instructions for $p01$, who was the only left-handed,
%were mistakenly given in a way that movements were performed different
%from what had been planned, and for participants $p13$ and $p16$ 
%data were corrupted because bluetooth communications problems with the 
%sensors. With that in mind, data for twenty participants were analysed in 
%this work.
%
%Of the twenty participants, all of them are right-handed healthy participants 
%of whom four are females and sixteen are males with a mean and standard 
%deviation (SD) age of mean=19.8 (SD=1.39). All participants provided 
%informed consent forms prior to participation in the experiment.
%
%\subsection*{Human-humanoid imitation activities}
%For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
%two of which were attached to the right hand of the participant and the 
%other two to the left hand of the humanoid robot.
%Then, each participant was asked to imitate repetitions of simple horizontal
%and vertical arm movements performed by the humanoid robot in the following 
%conditions:
%(i) ten repetitions of horizontal arm movement at normal (HN) 
%	and faster (HF) speed (Figure~\ref{fig:hri} A), and
%(ii) ten repetitions of vertical arm movement at normal (VN) 
%	and faster (VF) speed (Figure~\ref{fig:hri} C).
%The normal and faster speed of arm movements is defined by the duration 
%in number of samples of one repetition of NAO's arm movements. 
%We select NAO's arm movements duration to distinguish between normal and 
%faster arm movements as NAO's movements have less variation between 
%repetition to repetition. The duration for one repetition of the horizontal 
%arm movement at normal speed, HN, is about 5 seconds considering 
%that each repetition last around 250 samples.
%For horizontal arm movement at faster speed, HF, each repetition were performed 
%in around 2 seconds which correspond to 90 samples of data.
%The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
%which is around 300 samples of data. For vertical arm movement at 
%faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
%to 120 samples of data.
%To visualise the distinction between normal and faster speed for horizontal 
%and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
%for axes Z and Y of the gyroscope sensors with four window lengths: 
%2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
%and 15-sec (750-samples).
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht] 
%\centering
%\includegraphics[width=1.0\textwidth]{figures/experiment/speedtimeseries/v01/sts01}
%%\includegraphics[width=1.0\textwidth]{sts01}
%    \caption{
%	{\bf Time series duration of horizontal and vertical arm movements.} 
%    		Time series of smoothed data from gyroscope sensor 
%		for different speed arm movements performed by NAO: 
%		(A) Horizontal Normal arm movement (HN), 
%		(B) Horizontal Faster arm movement (HF),
%		(C) Vertical Normal arm movement (VN) and 
%		(D) Vertical Faster arm movement (VF).
%		Additionally, 
%		(A) shows window sizes for 2-seconds (100 samples), 
%		5-seconds (250 samples), 10-seconds (500 samples) 
%		and 15-seconds (750 samples)
%		which are also presented in (B), (C) and (D).
%		R code to reproduce figure is available \cite{hwum2018}.
%        }
%	\label{fig:sts}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%
%
%
%
%\subsection*{Data from Inertial Measurement Units} 
%	\label{sec:experiment:subsec:imu}
%\subsubsection*{Raw data}
%Considering the work of \cite{shoaib2016} which provided evidence of an 
%improvement in recognition activities when combining data from 
%accelerometer and gyroscope. We focus our analysis from data of the 
%accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and 
%leave the data of the magnetometer and quaternions for further investigation 
%because of their possible variations with regard to magnetic disturbances.
%
%Data from the accelerometer is defined by triaxial time series 
%$A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ 
%(Eq.~\ref{eq:A}), and the same for data from the gyroscope 
%which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ 
%representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:G}).
%Both triaxial time series of each sensor, $a$ and $g$, are denoted with 
%its respective axes subscripts $x,y,z$, where $n$ is the sample index 
%and $N$ is the same maximum length of all axes for the time series.
%Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
%%%---------------------------------(EQUATION)----------------------------------
%\begin{equation}\label{eq:A}
%\boldsymbol{A} =
%\begin{pmatrix}
%  A_x(n) \\
%  A_y(n) \\
%  A_z(n)
%\end{pmatrix}
%=
%\begin{pmatrix}
% a_x(1),a_x(2),\dots,a_x(N) \\
% a_y(1),a_y(2),\dots,a_y(N) \\
% a_z(1),a_z(2),\dots,a_z(N) 
%\end{pmatrix},
%\end{equation}
%%%---------------------------------(EQUATION)-----------------------------------
%and 
%%%---------------------------------(EQUATION)-----------------------------------
%\begin{equation}\label{eq:G}
%\boldsymbol{G} =
%\begin{pmatrix}
% G_x(n) \\
% G_y(n) \\
% G_z(n)
%\end{pmatrix}
%=
%\begin{pmatrix}
% g_x(1),g_x(2),\dots,g_x(N) \\
% g_y(1),g_y(2),\dots,g_y(N) \\
% g_z(1),g_z(2),\dots,g_z(N) 
%\end{pmatrix}.
%\end{equation}
%%%---------------------------------(EQUATION)-----------------------------------
%
%
%
%
%\subsubsection*{Postprocessing data}
%After the collection of raw data from four NeMEMsi sensors,
%time synchronisation alignment and interpolation were performed 
%in order to create time series with the same length and synchronised time.
%We refer the reader to \cite{Comotti2014} for further
%details about the time synchronisation process.
%
%\subsubsection*{Data normalization}
%Data is normalised to have zero mean and unit variance 
%using sample mean and sample standard deviation.
%The sample mean and sample standard deviation using $x(n)$ is given by
%%%---------------------------------(EQUATION)----------------------------------
%\begin{equation}\label{eq:ms}
%\mu_{x(n)}= \frac{1}{N} ( \sum_{i=1}^N x(i) ), \quad  \sigma_{x(n)} =  \sqrt{ \frac{  \sum_{1=1}^N ( x(i) - \mu_{x(n)} )^2 }{ N-1 }  },      
%\end{equation}
%%%---------------------------------(EQUATION)----------------------------------
%and the normalised data, $\hat{x}(n)$, is computed as follows
%%%---------------------------------(EQUATION)----------------------------------
%\begin{equation}\label{eq:normalization}
%\hat{x} (n) = \frac{   x(n) -  \mu_{x(n)}  }{   \sigma_{x(n)} }.   
%\end{equation}
%%%---------------------------------(EQUATION)----------------------------------
%
%\subsubsection*{Smoothing data}
%Commonly, a low-pass filter is the method either to capture the low 
%frequencies that represent \%99 of the human body energy or to get 
%the gravitational and body motion components of 
%accelerations \cite{anguita2013}. However, for this work the elimination of 
%certain range of frequencies is not the main focus but the conservation of 
%the structure in the time series in terms of the width and heights where, 
%for instance, Savitzky-Golay filter can help to accomplish 
%such task \cite{press1992}. Savitzky-Golay filter is based on the 
%principle of moving window averaging which preserves the area under 
%the curve (the zeroth moment) and its mean position in time 
%(the first moment) but the line width (the second moment) is violated 
%and that results, for example, in the case of spectrometric data
%where a narrow spectral line is presented with reduced height and width. 
%With that in mind, the aim of Savitzky-Golay filtering is to find filter 
%coefficients $c_n$ that preserve higher momentums which are based on local 
%least-square polynomial approximations \cite{savitzkygolay1964, 
%press1992, schafer2011}.
%Therefore, Savitzky-Golay coefficients are therefore computed using 
%an R function \texttt{sgolay(p,n,m)} where \texttt{p} is the filter order, 
%\texttt{n} is the filter length (must be odd) and \texttt{m} is the 
%$m$-th derivative of the filter coefficients  \cite{Rsignal}. 
%Smoothed signal is represented with a tilde over the original 
%signal: $\tilde{x}(n)$.
%
%\subsubsection*{Window size data}
%With regard to the window size, \cite{shoaib2016} investigated its effects 
%using seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds)
%and combination of inertial sensors (accelerometer, gyroscope and linear 
%acceleration sensor) to improve the activity recognition performance for 
%repetitive activities (walking, jogging and biking) and less repetitive 
%activities (smoking, eating, giving a talk or drinking a coffee).
%With that in mind, Shoaib et al. \cite{shoaib2016} concluded that the 
%increase of window length improve the recognition of complex activities 
%because these requires a large window length to learn the repetitive 
%motion patterns. Particularly, one of the recommendations is to use large 
%window size to recognise less repetitive activities which mainly involve 
%random hand gestures. Therefore, for the four activities 
%(HN, HF, VN, and VF) in this work, which are mainly repetitive, 
%we select only four window sizes for analysis: 2-s window (100 samples), 
%5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
%(Fig~\ref{fig:sts}).
%
%
%

%% As a guideline references should be limited to 60 (this is not 
%% strictly enforced).
%\bibliography{sample}
%\bibliography{references}
\bibliography{../references/references}

\section*{Acknowledgements}

This work was funded by the University of Birmingham and the 
Mexican National Council of Science and Technology as part of 
Miguel Xochicale's PhD degree. I would like to thank to Dr. Dolores 
Columba Perez Flores and Prof. Martin J Russell for their helpful 
comments to polish the use of the language of mathematics and to 
Constantino Antonio Garcia Martinez et al. for developing the R 
package nonlinearTseries that help to accelerate the 
analysis of the nonlinear time series in this work.

\section*{Author contributions statement}
Contribution for this work of 
Miguel Xochicale (MX) and Chris Baber (CB) is as follows:
\begin{description}
\item[Conceptualisation] MX, CB
\item[Data Curation] MX
\item[Formal Analysis] MX
\item[Funding Acquisition] MX, CB
\item[Investigation] MX
\item[Methodology] MX
\item[Project Administration] MX
\item[Resources] CB
\item[Software] MX
\item[Supervision] CB
\item[Validation] MX
\item[Verification] MX
\item[Writing - Original Draft Preparation] MX
\item[Writing - Review] MX, CB
\item[Writing - Editing] MX
\end{description}

%
%\section*{Additional information}
%
%To include, in this order: \textbf{Accession codes} 
%(where applicable); \textbf{Competing interests} (mandatory statement). 
%
%The corresponding author is responsible for submitting a 
%\href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on 
%behalf of all authors of the paper. This statement must be included 
%in the submitted article file.
%
%\begin{figure}[ht]
%\centering
%\includegraphics[width=\linewidth]{stream}
%\caption{Legend (350 words max). Example legend text.}
%\label{fig:stream}
%\end{figure}

%\begin{table}[ht]
%\centering
%\begin{tabular}{|l|l|l|}
%\hline
%Condition & n & p \\
%\hline
%A & 5 & 0.1 \\
%\hline
%B & 10 & 0.01 \\
%\hline
%\end{tabular}
%\caption{\label{tab:example}Legend (350 words max). Example legend text.}
%\end{table}
%
%Figures and tables can be referenced in LaTeX using the ref command, 
%e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.

\end{document}