### 2019-04-29 docs/corrections/29-03-2019/logbook/logbook.tex
[docs/corrections/29-03-2019] part-01

\title{Logbook for corrections }
\author{Miguel Xochicale}
\date{ \today }

\documentclass[12pt]{article}
\usepackage[margin=0.99in]{geometry}
\usepackage{enumitem}

% for asterik in circle
\usepackage{amsmath,amssymb,graphicx,color}
\newcommand{\invcircledast}{%
  \mathbin{\vphantom{\circledast}\text{%
    \ooalign{\smash{\blackcircle}\cr
             \hidewidth\smash{\textcolor{white}{$*$}}\hidewidth\cr
            }%
  }}%
}

\begin{document}
\maketitle

\begin{abstract}
This document presents a log-book for the corrections
of \emph{.../corrections/29-04-2019/comments/scanned-draft.pdf}.
\end{abstract}

%\tableofcontents
%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Title and authors}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}

	\title{The Application of Recurrent Quantification Analysis 
	to Human-Humanoid Interaction}

	Chris Baber

	\end{verbatim}
	\textit{
	SORTED:  Mon Apr  1 17:26:42 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Abstract}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$  Not a conclusion just an observation

	\begin{verbatim}

Then with the use of RQA metrics, we find out that the type of 
movements and the level of smoothness affects those metrics. 
Specifically, entropy values from RQA were well distributed and 
presented variation in all the conditions for time series. 
This work have then the potential to enhance the development of 
better diagnostic tools for various applications in rehabilitation, 
sport science or for new forms of human-robot interaction.

	\end{verbatim}
	\textit{
	SORTED:  Mon Apr  1 17:44:55 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
	added recommendations 	
	\end{verbatim}
	\textit{
	SORTED:  Wed Apr  3 18:08:38 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
	recommendations were added	
	\end{verbatim}
	\textit{
	SORTED:  Wed Apr  3 18:40:38 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
 	added recommendations	
	\end{verbatim}
	\textit{
	SORTED:  Thu Apr  4 12:46:47 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}

	\end{verbatim}
	\textit{
	SORTED:  
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}

	\end{verbatim}
	\textit{
	SORTED:  
	}
	\\

\end{enumerate}

\end{document}