### 2019-09-05 docs/corrections/29-03-2019/logbook/logbook.tex
completed corrections29-03-2019

\title{
	Logbook of corrections \\ 
	for manuscript-draft-29-04-2019 
}
\author{Miguel Xochicale}
\date{ \today }

\documentclass[12pt]{article}
\usepackage[margin=0.99in]{geometry}
\usepackage{enumitem}

% for asterik in circle
\usepackage{amsmath,amssymb,graphicx,color}
\newcommand{\invcircledast}{%
  \mathbin{\vphantom{\circledast}\text{%
    \ooalign{\smash{\blackcircle}\cr
             \hidewidth\smash{\textcolor{white}{$*$}}\hidewidth\cr
            }%
  }}%
}

\begin{document}
\maketitle

\begin{abstract}
This is a log book document for the corrections
of \emph{.../corrections/29-04-2019/comments/scanned-draft.pdf}.
\end{abstract}

%\tableofcontents
%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Title and authors}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}

	\title{The Application of Recurrent Quantification Analysis 
	to Human-Humanoid Interaction}

	Chris Baber

	\end{verbatim}
	\textit{
	SORTED:  Mon Apr  1 17:26:42 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Abstract}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$  Not a conclusion just an observation

	\begin{verbatim}

Then with the use of RQA metrics, we find out that the type of 
movements and the level of smoothness affects those metrics. 
Specifically, entropy values from RQA were well distributed and 
presented variation in all the conditions for time series. 
This work have then the potential to enhance the development of 
better diagnostic tools for various applications in rehabilitation, 
sport science or for new forms of human-robot interaction.

	\end{verbatim}
	\textit{
	SORTED:  Mon Apr  1 17:44:55 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
	added recommendations 	
	\end{verbatim}
	\textit{
	SORTED:  Wed Apr  3 18:08:38 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
	recommendations were added
	(i.e. types of signals were moved to methods section)	
	\end{verbatim}
	\textit{
	SORTED:  Wed Apr  3 18:40:38 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
 	added recommendations	
	\end{verbatim}
	\textit{
	SORTED:  Thu Apr  4 12:46:47 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
	Recommendations with the use of English language were added	
	\end{verbatim}
	\textit{
	SORTED:  Tue  3 Sep 07:15:23 BST 2019
	}
	\\

	\begin{verbatim}
	You need a sentence or two here
	to explain why this is claimed.
	\end{verbatim}
	\textit{
	SORTED:  
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
	Corrections with the use of English	
	\end{verbatim}
	\textit{
	SORTED:  Tue  3 Sep 07:26:41 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiment}

\begin{enumerate}

\item  (pp. )  \_$\wedge$

	\begin{verbatim}
	Needs a statment of ethics 
	Section of Ethics has been added. However I am not sure 
	if an Appendix should be added for further informotation.
	\end{verbatim}
	\textit{
	SORTED:  Thu  5 Sep 07:14:33 BST 2019
	}
	\\

\end{enumerate}

\end{document}