### 2020-07-02 docs/manuscript/main.tex

%*******************************************************************************
\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\graphicspath{{../}} %goes to path: docs/

\title{Recurrent Quantification Analysis of Movement Variability in Human-Humanoid Interaction Activities}
\author[1,*]{Miguel Xochicale}
\author[1]{Chris Baber}
\affil[1]{University of Birmingham, 
	School of Engineering, 
	Birmingham, 
	B15 2TT, 
	UK}
\affil[*]{map479@bham.ac.uk}

%\keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
Human movement variability occurs in motor performance 
across multiple repetitions of a task and such behaviour is an
inherent feature within and between each persons' movement. Quantifying
movement variability is still an open problem, particularly 
when methods in time domain, frequency domain or nonlinear dynamics 
can break down due to the real-world time-series datasets. 
For this work, we therefore investigate nonlinear dynamics methods 
such as reconstructed state space (RSSs), uniform time-delay embedding, 
recurrence plots (RPs) and recurrence quantification analysis metrics (RQAs)
with real-world time-series data.
Particularly, we are interested in the weaknesses and robustness 
of nonlinear dynamics tools with the use of raw and post-processed 
data of wearable inertial sensors. 
That said, twenty right-handed healthy participants 
imitated simple vertical and horizontal arm movements in normal 
and faster velocity from an humanoid robot in order to have 
four window lengths and three levels of smoothed time-series data,
to then found visual differences in the patterns with RSSs and RPs
and particulatly the computed differences with RQA metrics
that help us to quantify activities, types of sensors, windows lenghts 
and level of smoothness. 
Specifically, we can conclude that RQA ENTR, Shannon Entropy, 
can lead to interesting results on the quantification of 
movement variablity for participants of different ages, 
state of health and anthropomorphic features to 
then enhance the development of 
better diagnostic tools for applications in rehabilitation, 
sport science or for new forms of human-humanoid interaction.
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}
Human movement requires a complex system where not only multiple
joints and limbs are involved for a specific task in a determined environment
but also perception and action of movement that affects such 
physical performance \cite{davids2003}. 
In contrast, variability in humanoid movement is usually very small,
as a result of mechanical and dynamic capabilities \cite{gouaillier2009}. 
This means that, while humanoids 
solve the degrees of freedom problem through join design and algorithms,
humans tend to have more fluid and flexible approach. Consequently, 
one can see much variability in human movement performance of even the simplest task.
Studies of human motion reveal the possibility to 
estimate features from lower dimension signals to distinguish differences between 
styles of pedalling motion \cite{Quintana-Duque2012, Quintana-Duque2016}, 
gait identification \cite{sama2013, frank2010},
and detection of pathologies \cite{gomezgarcia2014}.
The lower dimension signals from biological systems are time series 
of $1-$dimension in $\mathbb{R}$ which commonly are 
noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
Hence, nonlinear dynamics can be used to 
objectively quantify variability of lower
dimension signals \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, 
frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
For instance, Bradley et al. 2015 \cite{bradley2015} reviewed methods for
nonlinear time series analysis based on the estimation of the 
embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to 
reconstruct the state space,
where an $n$-dimensional reconstructed state space using $1-$dimensional 
time series,
can preserve the topological properties of an unknown $M$-dimensional 
state space \cite{takens1981}.
Similarly, Bradley et al. 2015 \cite{bradley2015} reviewed the use of 
Recurrence Plots (RPs), a graphical representation of a two-dimensional map 
which show black and white dots as recurrences in a given $n$-dimensional system, 
and Recurrence Quantification Analysis (RQA) metrics that compute statistics in RPs.
In general, RPs and RQA provide an intuitive meaning of the time series,
for instance, RQA is quantitatively and qualitatively independent of 
embedding dimension (also verified experimentally \cite{iwanski1998}).
However, the estimation of embedding parameters and the selection of the 
right parameters to perform RQA is still an open problem.
Bradley et al. 2015 \cite{bradley2015} pointed out that
there is no general technique that can be used to compute the embedding
parameters since time series are system-dependent which means that
computing the values for embedding parameters may only work for
one purpose (e.g., prediction) and may not work well for another purpose
(e.g., computing dynamical invariants).
Additionally, methods of nonlinear dynamics for 
computing embedding parameters e.g., autocorrelation, mutual information, 
and nearest neighbour require data which is well 
sampled and with little noise \cite{garland2016}
or require signals that are purely deterministic \cite{kantz2003}.
Thus, these methods of nonlinear analysis for computing the embedding 
parameters can break down
with real-world datasets which have generally different length, 
different values of accuracy and precision (rounding errors due to finite 
precision of the measurement apparatus which include frequency 
acquisition \cite{frank2010}),
and contaminated data with different sources of noise
\cite{garland2016}.
It is then surprising that even with the previous constraints with regard to 
the quality of data, and the problem with the estimation of embedding parameters,
the use of nonlinear dynamics have proven to be helpful to understand and 
to characterise dynamics of time series 
\cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010,
gomezgarcia2014, gomezgarcia2014, marwan2011, stergiou2011, bradley2015}.

%Another point to consider with time series analysis using nonlinear dynamics
%is the appropriate use of different post-processing techniques such as 
%interpolation, filtering or normalisation.
%With that in mind, it can be said that there is little research and 
%understanding on the effects for post-processing techniques
%in the interpretation of reconstructed state spaces, RPs and RQA.

For this work, we are interested in exploring and investigating  
the effects of different features of time series 
(e.g. levels of smoothness, window data length,
structures of time series based on types of movements, 
types of sensors, participants and velocities) to estimate 
embedded parameters for uniform time-delay embedding, 
recurrence plots and recurrence quantification analysis.
Hence, we conducted an experiment with twenty right-handed 
healthy participants in the context of human-humanoid imitation 
activities where participants were asked to 
imitate simple arm movements performed by a humanoid robot.
The primary aim of this work is to explore the following questions:
\begin{itemize}
\item What are the effects on three nonlinear methods 
	(reconstructed state space with uniform time-delay embedding, 
	Recurrence Plots, and Recurrence Quantification Analysis), 
	for different embedding parameters, different recurrence thresholds 
	and different characteristics of time series 
	(structure of the signal, window length and smoothness of the signal)?, and 
\item What are the strengths and weaknesses of Recurrence
	Quantification Analysis for the previous conditions
	of nonlinear analysis methods?
\end{itemize}

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Results}

\subsection*{Reconstructed State Spaces}
One of the challenges of the implementation of 
uniform time-delay embedding to reconstruct the state spaces is the 
selection of embedding parameters because each time series is unique in 
terms of its structure (e.g. modulation of amplitude, frequency and phase) 
\cite{ frank2010, sama2013, bradley2015}.
With that in mind, the problem for this work is not to compute individual 
embedding parameters for each of the time series but to deal with a 
selection of two parameters that can represent all the time series. 
Our solution for such problem, as explained in below sections, 
was to compute a sample mean over all 
values in each of the conditions of the time series 
%of Figs~\ref{fig:ts}
for minimum dimension values and 
%Figs~\ref{fig:caoami} 
for minimum delay values.
%, resulting in an average minimum embedding parameters 
%of ($\overline{m_0}=6$, $\overline{\tau_0}=8$).

\subsubsection*{Minimum embedding parameters}
Minimum embedding parameters were firstly computed 
with the methods of False Nearest Neighbour and Average Mutual Information.
%for Uniform Time-Delay Embedding (UTDE),
%Reconstructed State Space (RSS), 
%Recurrence Plots (RPs) and 
%Recurrence Quantification Analysis (RQAs).
Figures \ref{fig:cao_ami}(A) show box plots for the minimum embedding values 
for sensors HS01 and RS01. Minimum embedding vaues for HS01 presents more 
variations than RS01 as interquartile range for HS01 is near to 1 (with the exceptions of HF sg0, VN sg0, and VF sg1)
whereas interquartile range for RS01 is near to 0.1 (with the exception of two axis HF sg0 and VF sg1).
Additionally, Figures \ref{fig:cao_ami}(A) show
a decrease of mean values (rhombus) in the box plots
as smoothness of time series increase (see sg0, sg1, sg2).
Figures \ref{fig:cao_ami}(B) show box plots for the first minimum values of AMI values.
It can be seen that values for HS01 tend to be more spread as the smoothness 
of the time series is increasing 
(see the increase of both mean (rhombus) and interquartile range).
However, AMI values for RS01 do not show such increase in relation with
the increase of smoothness except for HF and VF
(Figs \ref{fig:cao_ami}(B)).

We also computed a sample mean for an overall minimum embedding parameters that 
represent all participants, activities, sensors and levels of smoothness.
The sample mean for the minimum values of $E_{1}(m)$ from Figs \ref{fig:cao_ami}(A) 
is $\overline{m}_0=6$ 
and the sample mean for minimum values of AMIs from Figs~\ref{fig:cao_ami}(B)
is $\overline{\tau}_0=8$.
%Hence, we use the average minimum embedding parameters 
%($\overline{m_0}=6$, $\overline{\tau_0}=8$) to compute 
%Reconstructed State Spaces (RSSs), Recurrence Plots (RPs) and
%Recurrence Quantification Analysis (RQA) metrics for human-humanoid activities.
%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/caoami/pdf/fig2}
	\caption{
	{\bf Box plots of minimum embedding parameters.} 
	Box plots of (A) minimum embedding dimensions 
	and (B) first minimum AMI values for 
	Horizontal Normal (HN), Horizontal Faster (HF),
	Vertical Normal (VN) and Vertical Faster (VF)
	with sensors attached to participants (HS01) and
	sensor attached to robot (RS01).
	Minimum embedding parameters are for twenty participants 
	($p01$ to $p20$) with three smoothed signals 
	(sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ)
	and window length of 10-sec (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:cao_ami}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsubsection*{Uniform Time-Delay Embedding}
Uniform time-delay embedding were computed with the overall 
embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$) and 
the first three axis of the rotated data of the PCA are shown 
for the reconstructed state spaces of horizontal arm movements
(Figs~\ref{fig:rss_aHw10}) and vertical arm movements 
(Figs~\ref{fig:rss_aVw10}).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig3}
\caption{
	{\bf RSSs for horizontal arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters 
	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rss_aHw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig4}
    \caption{
	{\bf RSSs for vertical arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters
	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rss_aVw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

One can observe by eye the differences in each of the
trajectories of the reconstructed state spaces 
(Figs~\ref{fig:rss_aHw10} and \ref{fig:rss_aVw10}), 
however an objective quantification is required.
That being said, we tried to quantify those differences 
using euclidean distances between the origin 
to each of the points in their trajectories, 
which results did not capture the dynamics of 
the trajectories specially 
for those that looked very messy.
Such ambiguities lead us to consider the use of 
Recurrence Plots and Recurrence Quantification Analysis 
to objectively quantify dynamics the capture activities 
using time series.

\subsection*{Recurrences Plots}
Recurrence Plots (RP) were computed for horizontal arm movements (Fig~\ref{fig:rp_aH}) and 
vertical arm movements (Fig~\ref{fig:rp_aV}) 
using the average embedding parameters 
($\overline{m}_0=6$, $\overline{\tau}_0=8$) 
and an recurrence threshold of $\epsilon=1$.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aH}
\caption{
	{\bf RPs for horizontal arm movements.}	
	Recurrence plots were computed with 
	embedding parameters 
	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rp_aH}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aV}
\caption{
	{\bf RPs for vertical arm movements.}	
	Recurrence plots were computed with 
	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rp_aV}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsection*{Recurrence Quantification Analysis} \label{ch6:rqas}
Four Recurrence Quantification Analysis (RQA) metrics (REC, DET, RATIO and ENTR) 
were also computed with $\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.

\subsubsection*{REC values}
Figs~\ref{fig:RQABP}(A) show box plots of REC values 
that represent the \% of black dots in RPs.
It can be noted that REC values are more spread 
for HN and VN movements (higher interquartile range) than 
for HF and VF movements (lower interquartile range) for HS01 sensor. 
In contrast, REC values for RS01 sensor present little variation 
(interquartile range of 0.01).
Regarding the increase of smoothness for time series 
(sg0, sg1 and sg2), REC values present little 
variation as the smoothness is increasing for time series from HS01 
(changes of mean values (rhombus)) whereas REC values are 
more affected with the smoothness for data from RS01 
(see the incremental changes of mean values (rhombus)).
%See Figs~\ref{fig:rec_aH} and \ref{fig:rec_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual REC values for each participant.

\subsubsection*{DET values}
Figs \ref{fig:RQABP}(B) illustrate DET values 
that represent the predictability and organisation of RPs.
Generally, it can be noted little change of DET values 
(interquartile range is around 0.1) 
for type of movement, type of sensor 
but the increase of DET values as the smoothness of 
the signal increase 
(see the incremental changes of mean values (rhombus)).
However, the interquartile range for faster movements
(HF and VF) with no smoothing (sg0) is lower than the other
levels of smoothness (sg1 and sg2).
%See Figs~\ref{fig:det_aH} and \ref{fig:det_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual DET values for each participant.

\subsubsection*{RATIO values}
Figs \ref{fig:RQABP}(C) illustrate RATIO values that represent the 
dynamic transitions in RPs. 
Generally, it can be seen that RATIO values for HS01 sensor vary less 
for HN movements (interquartile range around 2)
than HF movements (interquartile range around 5).
For faster movements (HF and VF), it can be noted a decrease of 
the mean values the smoothness of the time series 
is increasing (rhombus). However, for normal movements (HN and VN), 
the decrease of mean values is less evident than 
faster movements (rhombus).
%See Figs~\ref{fig:ratio_aH} and \ref{fig:ratio_aV} in appendix 
%\ref{appendix:e:ep} 
%for more details about individual RATIO values for each participant.

\subsubsection*{ENTR values}
Figs \ref{fig:RQABP}(D) show ENTR values that represent 
Shannon entropy values in RPs. 
Generally, 
ENTR values for HS01 sensor show a higher variation  
(interquartile range around 0.5)
than ENTR values for RS01 sensor 
(interquartile range 0.1).
It can also be noted the increase of mean values 
for ENTR values as the smoothness of time series 
increase (rhombus) for type of movement, 
type of sensor and level of smoothness.  
%See Figs~\ref{fig:entr_aH} and \ref{fig:entr_aV} in appendix
%\ref{appendix:e:ep}
%for more details about individual ENTR values for each participant.

%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-bp}
    \caption
	[Box plots for RQA values]{
	{\bf Box plots for RQA metrics.}
	RQA metrics for (A) REC, (B) DET, (C) RATIO, and (D) ENTR of 
	20 participants performing HN, HF, VN and VF movements
	with sensors HS01, RS01 and three smoothed-normalised  
	time series (sg0, sg1 and sg2).
	RQA values were computed with 
	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$. 
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:RQABP}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

%\subsection*{RQA metrics with fixed parameters}
%Considering that RQA metrics were computed with fixed embedding parameters 
%($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found 
%the following. REC values, which represents the \% of black points in the RPs, 
%were more affected with and increase in normal speed movements (HN and VN) 
%than faster movements (HF and VF) for the sensor attached to the participants 
%(HS01). Such decrease of REC values from normal speed to faster speed 
%movements is also presented in data from sensor attached to the robot (RS01), 
%and little can be said with regard to the dynamics of the time series coming 
%from RS01 (Fig \ref{fig:RQABP}A).
%Similarly, DET values, representing predictability and 
%organisation in the RPs, present little variation in the any of the time 
%series where little can be said (Fig \ref{fig:RQABP}B).
%In contrast, RATIO values, which represent 
%dynamic transitions, were more variable for faster movements (HF and VF) 
%than normal speed movements (HN and VN) with sensors attached to the 
%participants (HS01). For data coming from sensors attached to the robot 
%(RS01), RATIO values from horizontal movements (HN, HF) appear to vary 
%more than values coming from vertical movmentes (VN, VF) 
%(Fig \ref{fig:RQABP}C).
%With that, it can be said that RATIO values can represent a bit better
%than REC or DET metrics for the variability of imitation activities in 
%each of the conditions for time series.
%Similarly, ENTR values for HN were higher than values for HF
%and ENTR values varied more for sensor attached to participants 
%than ENTR values for sensors of the robot. It is evidently that 
%the higher the entropy the more complex the dynamics are, 
%however, ENTR values for HN appear a bit higher than HF values, 
%for which we believe this happens because of the structure the time series
%which appear more complex for HN than  HF movements which presented a 
%more consistence repetition (Fig \ref{fig:RQABP}D).

\subsection*{3D RQA ENTR}
To show how that the selection of recurrence threshold affects RQA
values, 
we computed ENTR values of RQA metrics for different embedding parameters
$ \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
$  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ incrementing by one,
with the consideration of three recurrence thresholds $\epsilon=1, 2, 3$ 
and three levels of smoothness (sg0, sg1, sg2). 
For instance, 
Figures \ref{fig:RQA-IND} show the increase of recurrence threshold is 
associated to the increase of ENTR values in any of the levels 
of smoothness (see values of the ENTR bar).
Similarly, it can be noted that the increase of level of 
smoothness (sg0, sg1 and sg2) is associated with the increase 
of ENTR values of the 3D surface plots.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-epsilons}
    \caption
	[3D surface plots of RQA ENTR values]{
	{\bf 
	3D surface plots of RQA ENTR values for different recurrence threshold and smoothness levels.}
	RQA ENTR values are
	for embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \} $,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \} $
	incrementing by one and three recurrence thresholds $\epsilon=1, 2, 3$.
	RQA ENTR values were computed with data from $p03$, sensor HS01, with 
	a window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:RQA-IND}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
Figures~\ref{fig:3dRQAENTR_sensoractivities} show 3D surface plots of 
ENTR values for different sensors and different activities.
Figs~\ref{fig:3dRQAENTR_sensoractivities} show that RQA ENTR (lateral bars) 
decrease from normal (HN, VN) to faster (HF, VF) velocities from both 
human sensor (HS01) to robot sensor (RS01).
Also, it can be noted that RQA ENTR values decrease from human sensor (HS01) 
to robot sensor (RS01).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-sensors-activities}
    \caption{
	{\bf 3D surface plots of RQA ENTR values for different sensors and activities.}
	RQA ENTR values are for embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
	with $\epsilon = 1 $ considering four activities 
	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
	Vertical Faster (VF) and sensors Human Sensor 01 (HS01) and 
	Robot Sensor (RS01).
	RQA ENTR values were computed from data of $p03$, sg0 and 
	window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }
\label{fig:3dRQAENTR_sensoractivities}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------
Figures~\ref{fig:3dRQAENTR_participantsactivities} show the similarity of 3D surface plots in 
relationship with three participants (p01, p02 and p03)
performing four activities (HN, HF, VN, and VF) and two sensors (HS01, RS01).
Figures~\ref{fig:3dRQAENTR_participantsactivities}(A) 
present subtle differences for each of the participants (see RQA ENTR bar). 
Figures~\ref{fig:3dRQAENTR_participantsactivities}(B) show 
less change of RQA ENTR values for each participant as the data is 
from a sensor attached to the robot. 
Also, from data of sensors attached to the robot, it can be noted 
that changes are more notable for faster movements than normal movements. 
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\textwidth]{figures/rqa/output/rqa-participants}
    \caption{
	{\bf 3D surface plots of RQA ENTR values for different participants, activities and sensors.}
	RQA ENTR values are for participants (p01, p02, and p03) 
	in the categories of 
	(A) Human Sensor 01 (HS01) and 
	(B) Robot Sensor 01 (RS01)
	considering embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
	with $\epsilon = 1$ and four activities 
	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
	Vertical Faster (VF).
	RQA ENTR values were computed from data of sg0 and window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }
\label{fig:3dRQAENTR_participantsactivities}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------
Figures \ref{fig:3dRQAENTR_windowsactivities} show the effect of the increase 
of window length of the time series (w100 (2-sec), w250 (5-sec), w500 (10-sec) and w750 (15-sec))
on the 3D surface plots. 
It can be noted that the increase of number of samples improves the quality 
of the surface plots where 100 samples poorly capture the dynamics of each activity.
However as the window length increases from 250, 500 to 750 samples, 
3D surface plots show a similar representation of RQA ENTR values 
(being the surface plot with 750 samples the best test case).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-windows}
    \caption{
	{\bf 3D surface plots of RQA ENTR values for different windows lengths and activities.}
	RQA ENTR values are for embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$, 
	with $\epsilon = 1 $ considering four 
	windows lengths (e.g., w100 (100 samples), w250 (250 samples),
	w500 (500 samples) and w750 (750 samples)) and
	four activities 
	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
	Vertical Faster (VF).
	RQA ENTR values were computed from data of $p01$ and sg0.
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }
\label{fig:3dRQAENTR_windowsactivities}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

To summarise this section of results, it can be said that computing
embedding parameters for individual structure of time-series 
data is already a solved problem \cite{frank2010, sama2013, bradley2015}. 
However, it has been shown the challenge of finding embedding parameters 
for nonlinear dynamic tools that represent a set of different time-series data.
That said, we proposed the use of sample mean of the set of embedding parameters
for RSSs, RP and RQA to then noticed that the selection of recurrence 
threshold, $\epsilon$, is also an open problem.
For which, this work proposed the variation of recurrence thresholds 
and embedding parameters to show the relationships of these to different datasets 
(participants, activities, windows lengths and sensors).

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Discussion}
Time series from different sources  
(e.g., participants, movements, axis type) as well as
different characteristics (e.g., sample rate, window length or levels of smoothness) 
result in different embedding parameters and therefore different 
values for recurrence plots and recurrence quantification analysis metrics.
That said, while embedding parameters for individual
time series were successfully computed, the quantification 
of variability with regard to the shape of the trajectories in 
reconstructed state spaces require more investigation
as our initial, approach based on euclidean distances on the trajectories, 
failed to quantify trajectories which were not well unfolded. 
To then found out that Recurrence Plots along with 
Recurrence Quantification Analysis (RQA) metrics (e.g., REC, DET, RATIO and ENTR) 
and its variation of embedding parameters and recurrence thresholds 
helped to quantify the variability of different sources of time series.
However, further investigation is required to be done in order to 
have better intuition and meaningful interpretation of nonlinear 
analysis such as to find a right 
balance among (i) the level of smoothness of the signal, 
(ii) the selection of recurrence thresholds and (iii) 
the range of embedding parameters.

\section*{Conclusions}
This work allows to conclude that the choice of 
nonlinear analysis tool (e.g., RSSs, RPs, RQA metrics) will depend 
on what one would like to quantify on the time-series 
data (e.g., predicability, organisation, dynamics transitions, 
or complexity and determinism).
Then, time-series data characteristics 
(e.g., window size length, level of smoothness) plays an important role
as well on the results that nonlinear analysis tool can provide.
Similarly, the results of the nonlinear analysis tools
are associated with the structure of the time-series data 
(e.g. frequency, amplitude), the position of the sensor 
and activity performed by either a robot or human being
(as degrees of freedom from the humanoid are far less than
human movement). 
That said, it has been shown that the use of different 
characteristics of the time-series data 
(e.g., sensor, activity, level of smoothness and participants)
has help us to visualise and to quantify 
with nonlinear tools the variation of 
movements of, in this work, human-humanoid activities.
However, some limitation of nonlinear tools are related 
to the computation of different parameters 
(e.g., recurrence thresholds, embedding parameters)
that reflect the dynamics of individual characteristics 
of activity type, window length and structure of the time series.
Specifically, the example of DET values which appear 
to be constants across sensors, activities and levels of smoothness, 
whereas REC and RATIO, as function of REC, values 
show variation for certain sensors and movements. 
To then found out that RQA ENTR values with different recurrence 
thresholds were appropriate to quantify 
the different changes and variations of the characteristics of 
time-series data.
Therefore, we can conclude that this work provides a good starting 
point and reference to the use of Shannon Entropy to quantify 
human-humanoid imitation activities 
that can then lead to interesting results on the quantification 
of movement variability of participants with different ages, 
state of health and anthropomorphic features.

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Methods}
\subsection*{State Space Reconstruction}
The method of state space reconstruction \cite{packard1980} 
\cite{takens1981} has been applied in many disciplines 
\cite{aguirre2009, stergiou2011, frank2010, 
sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
based on uniform time-delay embedding methodology which is a simple 
matrix implementation that can reconstruct an unknown $d-$dimensional 
manifold $M$ from a scalar time series (e.g. one-dimensional 
time series in $\mathbb{R}$). A manifold, in this context, is a 
multidimensional curved surface within a space (e.g. a saddle) 
\cite{guastello-gregson2011}.

The use of a scalar time series is the main advantage of the method of state 
space reconstruction which in essence preserve dynamic invariants such as 
correlation dimension, fractal dimension, Lyapunov exponents, Kolmogorov-Sinai 
entropy and detrended fluctuation analysis \cite{bradley2015, 
Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
However, selecting appropriate embedding parameters which are sued to apply the 
state space reconstruction is still an open challenge for which we present 
introductions for the methodologies to compute such embedding parameters 
With that in mind, in the following subsections, we describe in more detail 
the state space reconstruction theorem (RSSs), uniform time-delay embedding 
theorem (UTDE), false nearest neighbours (FNN) and average mutual 
information (AMI). 
%and other methodologies for state space reconstruction.

\subsection*{State Space Reconstruction Theorem}
Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
%%********************************[EQUATION]************************************
\begin{equation}\label{eq:ssr}
  s(t)=f^t [s(0)],
\end{equation}
%%********************************[EQUATION]************************************
where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
iteration of $f$ that corresponds to an initial position 
$s(0) \in M $ \cite{takens1981}.
%$f$ is a smooth dynamical system, where smooth means a 
%continuously differentiable system
%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
can be obtained with
%%********************************[EQUATION]************************************
\begin{equation}\label{eq:measurement}
  x(t)=h[s(t)],
\end{equation}
%%********************************[EQUATION]************************************
where $h$ is a function, $h: M \rightarrow \mathbb{R}$, defined on the trajectory $s(t)$.
Reconstructed state space can then be described as an $n-$dimensional state 
space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
is the uniform time-delay embedding with a dimension embedding $m$
and delay embedding $\tau$ and
$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
of dimensionality (e.g. Principal Component Analysis, 
Singular Value Decomposition, etc) being $n \leq m$.
With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
$\boldsymbol{X}(t) = \Phi(s(t))$,
where $\Phi$ is a diffeomorphic map \cite{takens1981}
whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
dimension of $M$ \cite{garland2016}.
% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
% is called diffeomorphic if it is one-to-one correspondence and its inverse
% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
space then a composition of functions represented with $F^t$ is induced on the
reconstructed state space determined:
%by true smooth dynamical system $f^t$ and $\Phi$:
%********************************[EQUATION]************************************
\begin{equation}\label{eq:st}
  \boldsymbol{X}(t)=F^t [\boldsymbol{X}(0)] = \Phi \circ f^t \circ \Phi ^{-1}[\boldsymbol{X}(0)].
\end{equation}
%%********************************[EQUATION]************************************
With this in mind, an embedding is defined as "a smooth one-to-one 
coordinate transformation with a smooth inverse" and the total reconstruction 
map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
Fig~\ref{fig:ssr} illustrates the state space reconstruction.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/rss/pdf/rss}
    \caption{
	{\bf State space reconstruction methodology.}
	State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$
	where $f^t$ is the true dynamical system, $s(t)$ indicates the state,
    	$s$, at time, $t$,  and $h[ ]$ the measurement function. 
    	The time-delay embedding represented as the $\Phi$, maps the original
    	$d-$dimensional state $s(t)$ into the $m-$dimensional uniform 
	time-delay embedding matrix $\boldsymbol{X}(t)$.
    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into 
	a new state $y(t)$ of dimensions $n < m$.
    	(A) $M-$dimensional manifold representing the state space 
	(e.g. Lorenz system);
    	(B) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
   	(C) $m-$dimensional reconstructed state space with 
    	\texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and 
    	(D) $y(t)$ is the $n-$dimensional reconstructed state space.
	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $
	where $\Phi$ is the delay reconstruction map and 
	$\Psi$ is the coordinate transformation map.
	This figure is adapted from 
   	\cite{Quintana-Duque2012, casdagli1991, uzal2011}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
    }
    \label{fig:ssr}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

\subsection*{Uniform Time-Delay Embedding (UTDE)}\label{sec:utimedelayembedding}
Frank et al. and Sama et al. refer to the state space reconstruction
as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}. 
However, we consider the term "uniform time-delay embedding"
as more descriptive and appropriate terminology for our work.
Hence, the uniform time-delay embedding is represented as a matrix of uniform delayed 
copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is 
the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the 
samples of $\{ \boldsymbol{x}_n \}$.
$\{ \boldsymbol{x}_n \}_{n=1}^N$ has a sample rate of $T$.
The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$
and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$
where $i$ goes from $0,1, \dots, (m-1)$.
Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains
information of unobserved state variables and encapsulates the information of
the delayed copies of the available time series in the uniform time-delay 
embedding matrix $\boldsymbol{X}^{m}_{\tau}$, 
$\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
%%********************************[EQUATION]************************************
\begin{equation}\label{eq:tde}
\boldsymbol{X}^{m}_{\tau}  =
\begin{pmatrix}
\boldsymbol{ \tilde{x} }_n \\
\boldsymbol{ \tilde{x} }_{n-\tau} \\
\boldsymbol{ \tilde{x} }_{n-2\tau} \\
\vdots \\
\boldsymbol{ \tilde{x} }_{n- (m-1) \tau} \\
\end{pmatrix}^\intercal, 
\end{equation}
%%********************************[EQUATION]************************************
where $m$ is the embedding dimension, $\tau$ is the embedding delay and
$ ^\intercal$ denotes the transpose.
$m$ and $\tau$ are known as embedding parameters.
The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by
$N-(m-1)\tau$ rows and $m$ columns and 
$N-(m-1)\tau$ defines the length of each delayed copy 
of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
%For further details and explicit examples of uniform time-delay 
%embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.

\subsection*{Estimation of Embedding Parameters}
The estimation of the embedding parameters ($m$ and $\tau$) 
is a fundamental step for the state space reconstruction with the use
of uniform time-delay embedding method.
With this in mind, we review two of the most common algorithms,
which will be used in our work, to compute the embedding
parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%*******************************************************************************
\subsubsection*{False Nearest Neighbours}
To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992}
used the method of false neighbours which can be understood as follows:
on the one hand, when the embedding dimension is too small to unfold the 
attractor "not all points that lie close each other will be neighbours and 
some points appear as neighbours because of the attractor has been projected 
down into an smaller space", on the other hand, when increasing the embedding 
dimension "points that are near to each other in the sufficient
embedding dimension should remain close as the dimension increase from $m$ 
to $m+1$ \cite{krakovska2015}".
From a mathematical point of view, the state space reconstruction theorem is 
done when the attractor is unfolded with either the minimum embedding 
dimension, $m_0$, or any other embedding dimension value where 
$m \ge m_0$ \cite{kennel1992}. On the contrary, any large value of $m_0$ 
leads to excessive computations \cite{bradley2015}. With this in mind, 
Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method 
where only the time-series and one delay embedding value are necessary 
to select the minimum embedding dimension. Cao's algorithm is based 
on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
%%********************************[EQUATION]************************************
\begin{equation}\label{eq:e}
  \begin{aligned}
	E(m) &= \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau} a(i,m) \\
	 &=
       \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
       \frac{ || \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) || }
            { || \boldsymbol{X}_i(m) - \boldsymbol{X}_{n(i,m)}(m) ||  }
  \end{aligned}
\end{equation}
%%********************************[EQUATION]************************************
where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay
embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ 
and $\tau$ for which $E_1(m)$ is defined as
%%********************************[EQUATION]************************************
\begin{equation}\label{eq:e1}
E_1(m) = \frac{ E(m+1) } { E(m)}.
\end{equation}
%%********************************[EQUATION]************************************
$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$
in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater
than some $m_0$, if the time series comes from a multidimensional state space
then $m_0 + 1$ is the minimum dimension".
Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from
stochastic signals. $E_2(m)$ is defined as
%%********************************[EQUATION]************************************
\begin{equation}\label{eq:e2}
E_2(m) = \frac{ E^* (m+1) } { E^*(m)},
\end{equation}
%%********************************[EQUATION]************************************
where
%%********************************[EQUATION]************************************
\begin{equation}\label{eq:ee}
E^*(m) = \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
|  \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) |.
\end{equation}
%%********************************[EQUATION]************************************
For instance, when the signal comes from random noise (values that are 
independent from each other), all $E_2(m)$ values are approximately equal 
to 1 (e.g. $E_2(m) \approx 1$). However, for deterministic data $E_2(m)$ is 
not constant for all $m$ (e.g. $E_2(m) \neq 1$).

As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
series: the solution for the $x$ variable of the Lorenz system 
(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
and a variance of one (Fig~\ref{fig:e1e2}F).
We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
The $E_1(m)$ values for the chaotic time series appear to be constant
after the dimension is equal to six.
The determination of six is given that any value of $m$ can be used as they 
are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
$E_2(m)$ values, for chaotic time series, are different to one 
(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
chaotic time series the minimum embedding dimension the time series 
comes from a deterministic signal. With regard to the noise time series,  
$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
$E_1(m)$ values then indicate the minimum embedding dimension of the 
noisy time series is thirteen, however all of the $E_2(m)$ values are 
approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
concluded that noise time series is a stochastic signal.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/cao/pdf/cao}
    \caption{
	{\bf Minimum dimension embedding values with Cao's method.} 
	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
	with variations of $\tau$ values from one to twenty
	for (E) chaotic and (F) random time series.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:e1e2}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
computed but also a variation of $\tau$ from 1 to 20 is presented. 
The purpose of such variation for $\tau$ is to show its independence with
regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
(Fig~\ref{fig:e1e2}A,B,C, and D). 
However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
a new threshold where $m$ values appear to be constant in $E_1 (m)$.
In the case of the given examples and reported results, we defined such 
threshold as 0.05. Further investigation is required for the selection of the 
threshold in the $E_1(m)$, as the selection of the threshold in this work is 
base on no particular method but visual inspection.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%*******************************************************************************
\subsubsection*{Average Mutual Information}
When selecting the delay dimension parameter, $\tau$, one can consider the 
following two cases:
(i) when $\tau$ is too small, the elements of time-delay embedding will be along
the bisectrix of the phase space and the reconstruction is generally not 
satisfactory, 
(ii) on the contraty, when $\tau$ is too large the elements of the uniform 
time-delay embedding will become spread and uncorrelated which makes 
recovering the underlying attractor more difficult if not 
impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
With regard to the algorithms to compute $\tau$, 
Emrani et al. \cite{emrani2014a}, for instance, used the autocorrelation 
function in which the first zero crossing is considered as the minimum delay 
embedding parameter. However, the autocorrelation function is a linear 
statistic over which the Average Mutual Information (AMI) algorithm is 
preferred because the AMI takes into account the nonlinear dynamical 
correlations \cite{afraser1986,krakovska2015}. With this in mind, the AMI 
algorithm is described below to estimate the minimum delay embedding 
parameter, \texorpdfstring{$\tau_0$}{T}.

To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated
and then a probability distribution of data is computed \cite{kantz2003}.
AMI is therefore denoted by $I(\tau)$ which is the average mutual 
information between the original time series, $x(n)$, and the delayed 
time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}. 
AMI is defined by
%%********************************[EQUATION]************************************
\begin{equation}\label{eq:ami}
I(\tau) = \sum_{i,j}^N p_{ij} log_2 \frac{ p_{ij} }{ p_i p_j }.
\end{equation}
%%********************************[EQUATION]************************************
Probabilities are defined as follows:
$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of 
the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside 
the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability 
that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$. The AMI is measured 
in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
For small $\tau$, AMI will be large and it will then decrease more or 
less rapidly. As $\tau$ increase and goes to a large limit, 
$x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is 
factorised as $p_ip_j$ for which AMI is close to zero. Then, in order 
to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ 
where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or,
where the redundancy is the least" \cite{kantz2003}.

For example, we compute the AMI for two time series:
A) the $x$ solution of the Lorenz system, and
B) a noise time series using a normal distribution with mean zero and 
standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
concluded that the amount of knowledge for any noise time series is zero 
for which the first minimum embedding parameter is $\tau_0=1$. 
On the contrary, the first minimum of the AMI for the chaotic time series 
is $\tau_0=17$ which is the value that maximize the independence 
between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
\cite{bradley2015}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/ami/pdf/ami}
    \caption{
	{\bf Minimum delay embedding values with AMI's method.} 
    	(A, B) AMI values where its first minimum value in the curve
	is the minimum time delay embedding ($\tau_0$), 
	for (C) a chaotic and (D) noise time series.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:amis}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------
Similarly as Cao's algorithm negatives, AMI's algorithm is not an
exception for negatives, which are worthwhile to mention for further 
investigations.
For instance, (i) is not clear why the choose of the first minimum of the AMI 
is the minimum delay embedding parameter \cite{kantz2003} and 
(ii) the probability distribution of the AMI function is computed
with the use of histograms which depends on a heuristic choice of number of bins
for which AMI depends on partitioning \cite{garcia2005e71}.

%*******************************************************************************
\section*{Recurrence Quantification}\label{sec:recurrence-quantification}
\subsection*{Recurrence Plots}
Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences 
in conservative systems, however such discovery was not put into practice 
until the development of faster computers \cite{marwan2007},
for which Eckmann et al. \cite{eckmann1987} in 1987 introduced a method
where recurrences in the dynamics of a system can be visualised using 
Recurrence Plots. The intention of Eckmann et al. \cite{eckmann1987} was to 
propose a tool, called Recurrence Plot (RP), that provides insights into 
high-dimensional dynamical systems where trajectories are very difficult to 
visualise. Therefore, "RP helps us to investigate the 
$m-$dimensional phase space trajectories through a two-dimensional 
representation of its recurrences" \cite{marwan2015}.
Similarly, Marwan et al. \cite{marwan2015} pointed out that additionally 
to the methodologies of the state space reconstruction and other dynamic 
invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, 
the recurrences of the trajectories in the phase space can provide 
important clues to characterise natural process that present, for
instance, periodicities (as Milankovitch cycles) or irregular cycles 
(as El Ni\~no Southern Oscillation). 
Such recurrences can not only be presented visually using Recurrence Plots (RP) 
but also be quantified with Recurrence Quantification metrics, which leads 
to applications of these in various areas such as economy, physiology, 
neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.

For the creation of a recurrence plot based on time series 
$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
reconstruction with uniform time-delay embedding 
$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
The recurrence plot is therefore a two-dimensional $N \times N$ square 
matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
whenever $X(i)$ is sufficiently close to $X(j)$: 
%%********************************[EQUATION]************************************
\begin{equation}
\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
\end{equation}
%%********************************[EQUATION]************************************
where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
RP is also characterised with a line of identity (LOI) which is a  
diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/rp/pdf/rp}
    \caption{
	{\bf Recurrence Plots.} 
	(A) State space of the Lorenz system with controlling parameters 
	($\rho=28, \sigma=10, \beta=8/3$). 
	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
	(black circle) of a given point at $i$ is a recurrent point and is 
	represented as a black dot in the recurrence plot at location 
	$(i, j)$ or white otherwise.
	(B) Recurrence plot using the 
	three components of the Lorenz system and the RP with no embeddings 
	and threshold $\epsilon=5$.
	This figure is adapted from \cite{marwan2015}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
	}
    \label{fig:mrp}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\subsection*{Structures of Recurrence Plots}
Pattern formations in the RPs can be designated either 
as topology for large-scale patterns or texture for small-scale patterns.
In the case of topology, the following pattern formations are commonly presented:
(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
uniformly distributed noise (Fig~\ref{fig:rp2}A), 
(ii) periodic and quasi-periodic systems where diagonal lines and 
checkerboard structures represent oscillating systems, e.g., sinusoidal 
signals (Fig~\ref{fig:rp2}B), 
(iii) drift where paling or darkening recurrence points away from 
the LOI is caused by drifting systems, 
e.g., logistic map (Fig~\ref{fig:rp2}C), and
(iv) disrupted where recurrence points are presented white areas or 
bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
Texture patterns in RPs can be categorised as:
(i) single or isolated recurrence points that represent rare occurring states, 
and do not persist for any time or fluctuate heavily,
(ii) dots forming diagonal lines where the length of the small-scale parallel 
lines in the diagonal are related to the ratio of determinism or predictability 
in the dynamics of the system, and
(iii) dots forming vertical and horizontal lines where the length of the 
lines represent a time length where a state does not change or change very 
slowly and these patterns formation represent discontinuities in the signal, 
and (iv) dots clustering to inscribe rectangular regions which are related 
to laminar states or singularities \cite{marwan2015}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/rpsp/pdf/rpsp}
    \caption{
	{\bf Patterns in Recurrence Plots.} 
	Time-series with its respective recurrence plot for:
	(A) uniformly distributed noise,
	(B) super-positionet harmonic oscillation 
	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
	with a linearly increase term ($0.01*i$), and
	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
	This figure is adapted from \cite{marwan2015}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
	}
    \label{fig:rp2}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

Although, each of the previous pattern descriptions of the structures in the 
RP offer an idea of the characteristics of dynamical systems, 
these might be misinterpreted and conclusions might tend to be subjective 
as these require the interpretation of a particular researcher(s).
Because of that, recurrence quantification analyis (RQA) offer objective 
methodologies to quantify such visual characteristics of previous 
recurrent pattern structures in the RP \cite{zbilut1992}.

%*******************************************************************************
\subsection*{Recurrence Quantifications Analysis (RQA)}
Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
the density of recurrence points in RPs, then histograms of lengths for 
diagonal lines in RPs were studied by \cite{trulla1996} which were the 
introduction to the term recurrence quantification analysis (RQA) 
\cite{marwan2008}. RQA has been applied in many fields such as life science, 
engineering, physics, and others \cite{marwan2008}. Particularly in human 
movement to investigate noise and complexity of postural control 
\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
coordination \cite{duran2017}. The success of RQA is not only due to its 
simple algorithmic implementation but also to its capacity to detect tiny 
modulations in frequency or phase which are not detectable using standard 
methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
dimension which is verified experimentally by \cite{iwanski1998}.
RQA metrics comprehend percentage of recurrence, percentage of determinism, 
ratio, Shannon entropy of the frequency distributions of the line lengths,
maximal line length and divergence, trend and laminariy 
\cite{marwan2007, marwan2015}. For this work, we considered only four 
RQA metrics, due to its consistency with our preliminary experiments, 
which are described below. Such metrics are computed the nonlinearTseries 
R package \cite{nonlinearTseries2016}.

\subsubsection*{REC values}
The percentage of recurrence (REC) is defined as
%%********************************[EQUATION]************************************
\begin{equation}
REC(\epsilon,N) = 
		\frac{1}{N^2 - N} \sum^{N}_{i \neq j = 1} 
		\mathbf{R}^{m}_{i,j}(\epsilon),
\end{equation}
%%********************************[EQUATION]************************************
which enumerates the black dots in the RP excluding the line of identity.
REC is a measure of the relative density of recurrence points in the sparse 
matrix \cite{marwan2015}.
%REC is computed as follow with the nonlinearTseries package 
% \cite{nonlinearTseries} 
%  hist = getHistograms(neighs, ntakens, lmin, vmin)
%  # calculate the number of recurrence points from the recurrence rate. The
%  # recurrence rate counts the number of points at every distance in a concrete
%  # side of the main diagonal.
%  # Thus, sum all points for all distances, multiply by 2 (count both sides) and
%  # add the maindiagonal
%  numberRecurrencePoints = sum(hist$recurrenceHist) + ntakens
%  # calculate the recurrence rate dividing the number of recurrent points at a
%  # given distance by all points that could be at that distance
%  recurrence_rate_vector = 
%  hist$recurrenceHist[1:(ntakens - 1)] / ((ntakens - 1):1)
%  # percentage of recurrent points
%  REC = (numberRecurrencePoints) / ntakens ^ 2

\subsubsection*{DET values} 
The percent determinism (DET) is defined as the fraction of recurrence points
that form diagonal lines and it is determined by
%%********************************[EQUATION]************************************
\begin{equation}
DET=
	\frac{\sum^{N}_{l=d_{min}} l H_D{l} }
	     {\sum^{N}_{i,j=1} \mathbf{R}_{i,j}(\epsilon) },
\end{equation}
%%********************************[EQUATION]************************************
where 
%%********************************[EQUATION]************************************
\begin{equation}
H_D(l) = 
	\sum^{N}_{i,j=1} 
	(1- \mathbf{R}_{i-1,j-1}(\epsilon) ) 
	(1- \mathbf{R}_{i+l,j+l}(\epsilon) ) 
	\prod^{l-1}_{k=0}  \mathbf{R}_{i+k,j+k}(\epsilon)
\end{equation}
%%********************************[EQUATION]************************************
is the histogram of the lengths of the diagonal structures in the RP.
DET can be interpreted as the predictability of the system for periodic signals 
which, in essence, have longer diagonal lines than the short diagonals lines
for chaotic signals or absent diagonal lines for stochastic signals 
\cite{marwan2007, marwan2015}. Similarly, DET is considered as a measurement for 
the organisation of points in RPs  \cite{iwanski1998}. 
%percent determinism (DET) is computed as follow with the nonlinearTseries 
%package \cite{nonlinearTseries}  
% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
%                                       lmin = 2, lDiagonalHistogram,
%                                       recurrence_rate_vector, maxDistanceMD) {
%  #begin parameter computations
%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
%  DET = num / numberRecurrencePoints

\subsubsection*{RATIO values}
RATIO is defined as the ratio between DET and REC and it is calculated from 
the frequency distributions of the lengths of the diagonal lines.
RATIO is useful to discover dynamic transitions \cite{marwan2015}.
%  diagP = calculateDiagonalParameters(
%    ntakens, numberRecurrencePoints, lmin, hist$diagonalHist,
%    recurrence_rate_vector, maxDistanceMD
%  )
% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
%                                       lmin = 2, lDiagonalHistogram,
%                                       recurrence_rate_vector, maxDistanceMD) {
%  #begin parameter computations
%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
%  DET = num / numberRecurrencePoints
% 
%
%    RATIO = diagP$DET / REC
%

\subsubsection*{ENT values}
ENT is the Shannon entropy of the frequency distribution of the diagonal 
line lengths and it is defined as
%%********************************[EQUATION]************************************
\begin{equation}
ENT= - \sum^{N}_{l=d_{min}} p(l) ln p(l) \quad with 
	\quad p(l)=\frac{ H_D(l) }{ \sum^{N}_{ l=d_{min} } H_D(l) }.
\end{equation}
%%********************************[EQUATION]************************************
ENT reflects the complexity of the deterministic structure in the system.
For instance, for uncorrelated noise or oscillations, 
the value of ENT is rather small and indicates low complexity of the system,
therefore "the higher the ENT is the more complex the dynamics are" 
\cite{marwan2007, marwan2015}.
%#'  \item \emph{ENTR}: Shannon entropy of the diagonal line lengths distribution
%
%calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
%                                       lmin = 2, lDiagonalHistogram,
%                                       recurrence_rate_vector, maxDistanceMD) {
%
%  pl = lDiagonalHistogram / sum(lDiagonalHistogram)
%  diff_0 = which(pl > 0)
%  ENTR = -sum(pl[diff_0] * log(pl[diff_0]))

\subsection*{Sensitivity and robustness of RPs and RQA.}
RP and RQA are a very young field in nonlinear dynamics and many questions 
are still open, for instance, different parameters for window length size 
of the time series, embedding parameters or recurrence threshold can 
generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.

The selection of recurrence threshold, $\epsilon$, can depend on the system 
that is analysed. For instance, when studying dynamical invariants $\epsilon$ 
require to be very small, for trajectory reconstruction $\epsilon$ requires 
to have a large thresholds or when studying dynamical transition 
there is little importance about the selection of the threshold 
\cite{marwan2011}. Other criteria for the selection of $\epsilon$ is that 
the recurrence threshold  should be five times larger 
than the standard deviation of the observational noise
or the use of diagonal structures within the RP is suggested in order
to find the optimal recurrence threshold for (quasi-)periodic process 
\cite{marwan2011}.
Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
of choosing the right embedding parameters to perform RQA for which 
many experiments have to be performed using different parameters in order 
to have a better intuition of the nature of the time series and how 
this is represented by using RQA.

With that in mind, this work explores the sensitivity and robustness 
of the window size of time series, embedding parameters for RSS with UTDE 
and recurrence threshold for RP and RQA in order to gain a better insight 
into the underlying time series collected from inertial sensors in the 
context of human-humanoid imitation activities.
%Such changes of recurrence threshold values 
%can modify the patterns in RPs and therefore the values of RQA metrics.
%We therefore computed 3D surfaces to explore the sensibility and robustness of 
%embedding parameters and recurrence threshold in RQA  metrics. Following the 
%same methodology of computing 3D surfaces, we also considered variation of 
%window length size to present RQA metrics dependencies with embedding 
%parameters, recurrence thresholds and window length size.

%Choosing an appropriate recurrence threshold is crucial so as to get 
%meaningful representations in RP and RQA, however,

%Iwansky et al. \cite{iwanski1998} stated that patterns in Recurrence Plots and 
%metrics for RQA are independent of embedding dimension parameters, 
%however, that is not the case when using 
%different recurrence thresholds.

%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
%determining embedding parameters, their method consisted on creating 3D 
%surfaces with RQA metrics with an increase of embedding parameters 
%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
%fluctuations and gradual changes in the 3D surfaces that provide information 
%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
%created 3D surfaces for visual selection of not only embedding parameters 
%but also recurrence thresholds. Following same methodologies, we explored 
%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
%using 3D surfaces by an unitary increase of the pair embedding 
%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
%(Fig.~\ref{fig:topo_rqas}).

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Experiment} \label{sec:experiment}
We conducted an experiment in the context of human-humanoid imitation (HHI) 
activities where participants were asked to imitate simple horizontal and 
vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
Such simple movements were repeated ten times for the participant 
who copied NAO's arm movements in a face-to-face imitation activity.
Also, wearable inertial measurement unit (IMU) sensors were attached 
to the right hand of the participant and to the left hand of the robot 
(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
accelerometer, gyroscope and magnetometer sensors and quaternions 
\cite{Comotti2014}.

\subsection*{Ethics}
The experiments of this work were conducted in November 2016 and participants
confirmed reading and understanding the participant information sheet of the 
experiments and were able to withdraw from the experiment at any time without giving
any reason. The design of the experiments adhered to the University of X
regulations, data were anonymised and videos were stored only on a 
personal computer in accordance with the Data Protection Act 1998. 
%Refer to Appendix D for further information about the ethics, 
%online participation information sheets and experiment check list.

%A further description of the NeMEMSi IMU sensors 
%is given in \nameref{S1_AppendixB}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
  \centering
\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
    \caption{
	{\bf Human-humanoid imitation activities.} 
    		Face-to-face human-humanoid imitation (HHI) activities for 
		(A) HHI of horizontal arm movement, 
		(B) Humanoid horizontal arm movement,
		(C) HHI of vertical arm movement, and 
		(D) Humanoid vertical arm movement.
        }
    \label{fig:hri}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsection*{Participants}
Twenty-three participants,
from now on defined as $pN$ where $N$ is the number of participant, were 
invited to do the experiment. However, data for three participants were 
not used because the instructions for $p01$, who was the only left-handed,
were mistakenly given in a way that movements were performed different
from what had been planned, and for participants $p13$ and $p16$ 
data were corrupted because bluetooth communications problems with the 
sensors. With that in mind, data for twenty participants were analysed in 
this work.

Of the twenty participants, all of them are right-handed healthy participants 
of whom four are females and sixteen are males with a mean and standard 
deviation (SD) age of mean=19.8 (SD=1.39). All participants provided 
informed consent forms prior to participation in the experiment.

\subsection*{Human-humanoid imitation activities}
For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
two of which were attached to the right hand of the participant and the 
other two to the left hand of the humanoid robot.
Then, each participant was asked to imitate repetitions of simple horizontal
and vertical arm movements performed by the humanoid robot in the following 
conditions:
(i) ten repetitions of horizontal arm movement at normal (HN) 
	and faster (HF) speed (Figure~\ref{fig:hri} A), and
(ii) ten repetitions of vertical arm movement at normal (VN) 
	and faster (VF) speed (Figure~\ref{fig:hri} C).
The normal and faster speed of arm movements is defined by the duration 
in number of samples of one repetition of NAO's arm movements. 
We select NAO's arm movements duration to distinguish between normal and 
faster arm movements as NAO's movements have less variation between 
repetition to repetition. The duration for one repetition of the horizontal 
arm movement at normal speed, HN, is about 5 seconds considering 
that each repetition last around 250 samples.
For horizontal arm movement at faster speed, HF, each repetition were performed 
in around 2 seconds which correspond to 90 samples of data.
The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
which is around 300 samples of data. For vertical arm movement at 
faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
to 120 samples of data.
To visualise the distinction between normal and faster speed for horizontal 
and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
for axes Z and Y of the gyroscope sensors with four window lengths: 
2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
and 15-sec (750-samples).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht] 
\centering
\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
    \caption{
	{\bf Time series duration of horizontal and vertical arm movements.} 
    		Time series of smoothed data from gyroscope sensor 
		for different speed arm movements performed by NAO: 
		(A) Horizontal Normal arm movement (HN), 
		(B) Horizontal Faster arm movement (HF),
		(C) Vertical Normal arm movement (VN) and 
		(D) Vertical Faster arm movement (VF).
		Additionally, 
		(A) shows window sizes for 2-seconds (100 samples), 
		5-seconds (250 samples), 10-seconds (500 samples) 
		and 15-seconds (750 samples)
		which are also presented in (B), (C) and (D).
		Code and data to reproduce the figure is available in \cite{srep2020}.
        }
	\label{fig:sts}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsection*{Data from Inertial Measurement Units} 
	\label{sec:experiment:subsec:imu}

To give insight to the research questions, 
we considered various conditions of time series collected for this work 
(see Experiment section for more details) which are described as follows
\begin{itemize}
\item Three levels of smoothness for the normalised data 
	(sg0zmuv, sg1zmuv and sg2zmuv), computed from two different filter 
	lengths (29 and 159) with the same polynomial degree 
	of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal},
\item four velocities arm movement activities: 
	horizontal normal (HN), horizontal faster (HF), 
	vertical normal (VN), and vertical faster (VF), and
\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
	10-sec (500 samples) and 15-sec (750 samples).
\end{itemize}
%\subsection*{Time series}
After the data collection, raw time series were windowed, normalised and 
smoothed. Then, due to space limitations and to have simple visualisation, 
we only present 10-sec (500 samples) window length time series for 
three participants (p01, p01 and p03) performing horizontal 
arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}). 
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
    	\caption{ 
	{\bf Time series for horizontal and vertical arm movements.}
		(A/D) raw-normalised (sg0zmuv), 
		(B/E) normalised-smoothed 1 (sg1zmuv) and
		(C/F) normalised-smoothed 2 (sg2zmuv).
		Time series are only for three participants 
		($p01$, $p02$, and $p03$) 
		for horizontal and vertical arm movements in normal 
		and faster velocity (HN, HF, VN, VF) 
		with the normalised GyroZ or GyroY axis 
		and with one sensor attached to the participant (HS01) 
		and other sensor attached to the robot (RS01).	
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:ts}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsubsection*{Raw data}
Considering the work of \cite{shoaib2016} which provided evidence of an 
improvement in recognition activities when combining data from 
accelerometer and gyroscope. We focus our analysis from data of the 
accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and 
leave the data of the magnetometer and quaternions for further investigation 
because of their possible variations with regard to magnetic disturbances.

Data from the accelerometer is defined by triaxial time series 
$A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ 
(Eq.~\ref{eq:A}), and the same for data from the gyroscope 
which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ 
representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:G}).
Both triaxial time series of each sensor, $a$ and $g$, are denoted with 
its respective axes subscripts $x,y,z$, where $n$ is the sample index 
and $N$ is the same maximum length of all axes for the time series.
Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
%%---------------------------------(EQUATION)----------------------------------
\begin{equation}\label{eq:A}
\boldsymbol{A} =
\begin{pmatrix}
  A_x(n) \\
  A_y(n) \\
  A_z(n)
\end{pmatrix}
=
\begin{pmatrix}
 a_x(1),a_x(2),\dots,a_x(N) \\
 a_y(1),a_y(2),\dots,a_y(N) \\
 a_z(1),a_z(2),\dots,a_z(N) 
\end{pmatrix},
\end{equation}
%%---------------------------------(EQUATION)-----------------------------------
and 
%%---------------------------------(EQUATION)-----------------------------------
\begin{equation}\label{eq:G}
\boldsymbol{G} =
\begin{pmatrix}
 G_x(n) \\
 G_y(n) \\
 G_z(n)
\end{pmatrix}
=
\begin{pmatrix}
 g_x(1),g_x(2),\dots,g_x(N) \\
 g_y(1),g_y(2),\dots,g_y(N) \\
 g_z(1),g_z(2),\dots,g_z(N) 
\end{pmatrix}.
\end{equation}
%%---------------------------------(EQUATION)-----------------------------------

\subsubsection*{Postprocessing data}
After the collection of raw data from four NeMEMsi sensors,
time synchronisation alignment and interpolation were performed 
in order to create time series with the same length and synchronised time.
We refer the reader to \cite{Comotti2014} for further
details about the time synchronisation process.

\subsubsection*{Data normalization}
Data is normalised to have zero mean and unit variance 
using sample mean and sample standard deviation.
The sample mean and sample standard deviation using $x(n)$ is given by
%%---------------------------------(EQUATION)----------------------------------
\begin{equation}\label{eq:ms}
\mu_{x(n)}= \frac{1}{N} ( \sum_{i=1}^N x(i) ), \quad 
	\sigma_{x(n)} =  \sqrt{ \frac{  \sum_{1=1}^N ( x(i) - \mu_{x(n)} )^2 }{ N-1 }  },      
\end{equation}
%%---------------------------------(EQUATION)----------------------------------
and the normalised data, $\hat{x}(n)$, is computed as follows
%%---------------------------------(EQUATION)----------------------------------
\begin{equation}\label{eq:normalization}
\hat{x} (n) = \frac{   x(n) -  \mu_{x(n)}  }{   \sigma_{x(n)} }.   
\end{equation}
%%---------------------------------(EQUATION)----------------------------------

\subsubsection*{Smoothing data}
Commonly, a low-pass filter is the method either to capture the low 
frequencies that represent \%99 of the human body energy or to get 
the gravitational and body motion components of 
accelerations \cite{anguita2013}. However, for this work the elimination of 
certain range of frequencies is not the main focus but the conservation of 
the structure in the time series in terms of the width and heights where, 
for instance, Savitzky-Golay filter can help to accomplish 
such task \cite{press1992}. Savitzky-Golay filter is based on the 
principle of moving window averaging which preserves the area under 
the curve (the zeroth moment) and its mean position in time 
(the first moment) but the line width (the second moment) is violated 
and that results, for example, in the case of spectrometric data
where a narrow spectral line is presented with reduced height and width. 
With that in mind, the aim of Savitzky-Golay filtering is to find filter 
coefficients $c_n$ that preserve higher momentums which are based on local 
least-square polynomial approximations \cite{savitzkygolay1964, 
press1992, schafer2011}.
Therefore, Savitzky-Golay coefficients are therefore computed using 
an R function \texttt{sgolay(p,n,m)} where \texttt{p} is the filter order, 
\texttt{n} is the filter length (must be odd) and \texttt{m} is the 
$m$-th derivative of the filter coefficients  \cite{Rsignal}. 
Smoothed signal is represented with a tilde over the original 
signal: $\tilde{x}(n)$.

\subsubsection*{Window size data}
With regard to the window size, \cite{shoaib2016} investigated its effects 
using seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds)
and combination of inertial sensors (accelerometer, gyroscope and linear 
acceleration sensor) to improve the activity recognition performance for 
repetitive activities (walking, jogging and biking) and less repetitive 
activities (smoking, eating, giving a talk or drinking a coffee).
With that in mind, Shoaib et al. \cite{shoaib2016} concluded that the 
increase of window length improve the recognition of complex activities 
because these requires a large window length to learn the repetitive 
motion patterns. Particularly, one of the recommendations is to use large 
window size to recognise less repetitive activities which mainly involve 
random hand gestures. Therefore, for the four activities 
(HN, HF, VN, and VF) in this work, which are mainly repetitive, 
we select only four window sizes for analysis: 2-s window (100 samples), 
5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
(Fig~\ref{fig:sts}).

%% As a guideline references should be limited to 60 (this is not 
%% strictly enforced).
%\bibliography{sample}
%\bibliography{references}
\bibliography{../references/references}

\section*{Acknowledgements}

This work was funded by the University of Birmingham and the 
Mexican National Council of Science and Technology as part of 
Miguel Xochicale's PhD degree. I would like to thank to Dr. Dolores 
Columba Perez Flores and Prof. Martin J Russell for their helpful 
comments to polish the use of the language of mathematics and to 
Constantino Antonio Garcia Martinez et al. for developing the R 
package nonlinearTseries that help to accelerate the 
analysis of the nonlinear time series in this work.

\section*{Author contributions statement}
Contributions for this work of 
Miguel Xochicale (MX) and Chris Baber (CB) are as follows:
\begin{description}
\item[Conceptualisation] MX, CB
\item[Data Curation] MX
\item[Formal Analysis] MX
\item[Funding Acquisition] MX, CB
\item[Investigation] MX
\item[Methodology] MX
\item[Project Administration] MX
\item[Resources] CB
\item[Software] MX
\item[Supervision] CB
\item[Validation] MX
\item[Verification] MX
\item[Writing - Original Draft Preparation] MX
\item[Writing - Review] MX, CB
\item[Writing - Editing] MX
\end{description}

%
%\section*{Additional information}
%
%To include, in this order: \textbf{Accession codes} 
%(where applicable); \textbf{Competing interests} (mandatory statement). 
%
%The corresponding author is responsible for submitting a 
%\href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on 
%behalf of all authors of the paper. This statement must be included 
%in the submitted article file.
%
%\begin{figure}[ht]
%\centering
%\includegraphics[width=\linewidth]{stream}
%\caption{Legend (350 words max). Example legend text.}
%\label{fig:stream}
%\end{figure}

%\begin{table}[ht]
%\centering
%\begin{tabular}{|l|l|l|}
%\hline
%Condition & n & p \\
%\hline
%A & 5 & 0.1 \\
%\hline
%B & 10 & 0.01 \\
%\hline
%\end{tabular}
%\caption{\label{tab:example}Legend (350 words max). Example legend text.}
%\end{table}
%
%Figures and tables can be referenced in LaTeX using the ref command, 
%e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.

\end{document}